### 2018-10-16 paper.tex
Initial ShareLaTeX Import

\documentclass[a4paper,10pt]{article}
% * <m.f.schepers@rug.nl> 2018-10-12T08:13:57.409Z:
%
% ^.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%											Packages											  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{enumerate}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{bm}
\usepackage[mathscr]{eucal}
\usepackage{a4wide}
\usepackage{authblk}
\usepackage{appendix}
\usepackage{hyperref}
\hypersetup{
    colorlinks = true,
    urlcolor = {magenta},
    citecolor = {green},
    linkcolor = {blue}
}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,backgrounds,patterns,decorations.markings,decorations.pathreplacing,intersections}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}

\usepackage{subcaption}
\usepackage{dsfont}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%											Macros												  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                        Probability                                              %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\pr}{\mathbb{P}}								%	short hand for blackboard P
\newcommand{\Pee}{\mathbb{P}}

%% Probability commands %%

\newcommand{\Prob}[1]{\pr\left(#1\right)}					%	Standard probability command, 
															%	argument is event

\newcommand{\Probn}[1]{\pr_n\left(#1\right)}				%	Probability command with subscript n

\newcommand{\Probc}[2]{\pr_{#2}\left(#1\right)}				%	Probability command conditioned on 
															%	the points of the graph specified in
															%	the second argument

\newcommand{\CProb}[2]{\pr\left(\left.#1\right|#2\right)}	%	Conditional probability command, 
															%	second argument is conditioned 
															%	event/random variable

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                         Expectation                                             %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\e}{\mathbb{E}}								%	short hand for blackboard E
\newcommand{\E}{\mathbb{E}}

%% Expectation commands %%

\newcommand{\Exp}[1]{\e\left[#1\right]}					%	Standard expectation command, argument
														%	is random variable

\newcommand{\Expn}[1]{\e_n\left[#1\right]}				%	Expectation with subscript n

\newcommand{\Expc}[2]{\e_{#2}\left[#1\right]}			%	Expectation command conditioned on 
														%	the points of the graph specified in
														%	the second argument

\newcommand{\CExp}[2]{\e\left[\left.#1\right|#2\right]}	%	Conditional expectation, second 
														%	argument is conditioned event/random 
														%	variable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     Probabilistic limits                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\plim}{\ensuremath{\stackrel{\pr}{\rightarrow}}}	%	Convergence in probability
\newcommand{\dlim}{\ensuremath{\stackrel{d}{\rightarrow}}}		%	Convergence in distribution

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                        Scaling notations                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\bigO}[1]{O\left(#1\right)}				%	Big O notation
\newcommand{\smallO}[1]{o\left(#1\right)}			%	Small o notation
\newcommand{\bigT}[1]{\Theta\left(#1\right)}		%	Theta notation
\newcommand{\bigOp}[1]{O_\pr \left(#1\right)}		%	Probabilistic big O notation
\newcommand{\smallOp}[1]{o_\pr \left(#1\right)}		%	Probabilistic little o notation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                        Indicator commands                                       $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\ind}[1]{\mathbbm{1}_{\left\{#1\right\}}}	%	Command for indicator where argument is a condition
\newcommand{\indE}[1]{\mathbbm{1}_{#1}}					%	Command for indicator where the argument is an event 
\newcommand{\1}{\mathds{1}}								%	Indication shorthand command

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                       Additional commands                                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\N}{\mathbb{N}}							%	Natural numbers
\newcommand{\R}{{\ensuremath{\mathbb R}}}			%	Real numbers

%%	Short hand commands:

\newcommand{\Kcal}{\ensuremath{{\mathcal K}}}		%	Mathcal K
\newcommand{\Bcal}{\ensuremath{{\mathcal B}}}		%	Mathcal B
\newcommand{\Ccal}{\ensuremath{{\mathcal C}}}		%	Mathcal C
\newcommand{\Ncal}{\ensuremath{{\mathcal N}}}		%	Mathcal N

\newcommand{\eps}{\varepsilon}						%	varepsilon
\DeclareMathOperator{\dd}{\, \mathrm{d}\hspace{-1pt}}			%	straight d for integration

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                     Paper specific commands                                     %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\renewcommand{\H}{\ensuremath{\mathbb{H}}}			%	H used for the hyperbolic plane and hyperbolic random graph
\newcommand{\HP}{\ensuremath{\widetilde{\H}}}		%	Poisson hyperbolic random graph
\newcommand{\Dcal}[0]{\ensuremath{{\mathcal D}}}	%	Hyperbolic disc
\newcommand{\Rcal}[0]{\ensuremath{{\mathcal R}}}	%	Image of the hyperbolic disc under \Psi
\newcommand{\Pcal}[0]{\ensuremath{{\mathcal P}}}	%	Point process
\newcommand{\Po}[0]{\ensuremath{{\mathrm{Po}}}}		%	Poisson random variable

\newcommand{\T}{\ensuremath{T}}						%	Symbol for triangles

\newcommand{\BallHyp}[1]{\Bcal_{\H,n}\left(#1\right)}		%	The image of the hyperbolic ball around \Psi^{-1}(p)
\newcommand{\BallPo}[1]{\Bcal_{\Pcal}\left(#1\right)}		%	The neighborhood of a point p in the infinite limit model
\newcommand{\BallPon}[1]{\Bcal_{\Pcal,n}\left(#1\right)}	%	The neighborhood of a point p in the finit box model
\newcommand{\FatBallHyp}[1]{\check{\Bcal}_{\H,n}(#1)}		%
\newcommand{\BallSym}[1]{\Bcal_{\H \bigtriangleup \Pcal}\left(#1\right)}	%	Symmetric difference between balls in the 
																			%	hyperbolic and infinite limit model
\newcommand{\BallInter}[1]{\Bcal_{\H \cap \Pcal}\left(#1\right)}

\newcommand{\Mu}[1]{\mu_{\alpha,\nu}\left(#1\right)}

\newcommand{\hsym}[1]{h_{\H \bigtriangleup \Pcal}(#1)}
\newcommand{\EdgeDiff}{E_{\H \setminus \mathcal{P}}}
\newcommand{\expH}{\overline{n}_{\H}(k_n)}
\newcommand{\expP}{\overline{n}_{\Pcal}(k_n)}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}	%	Use to insert equation number 
																		%	inside * environment

\newcommand{\MeijerG}[7]{G_{#3,#4}^{#1,#2} \left(\begin{matrix}#5  \\ #6 \end{matrix} \bigg| #7 \right)}
\newcommand{\MeijerGnew}[7]{G_{#3,#4}^{#1,#2} \left(#7 \bigg|\begin{matrix}#5  \\ #6 \end{matrix}\right)}

\newcommand{\Real}{\operatorname{Re}}
\newcommand{\Li}{\operatorname{Li}}

\newcommand{\Ee}[0]{\ensuremath{{\mathbb E}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%								For leaving comments to each other:								  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\definecolor{orange}{RGB}{255,127,0}
\definecolor{pink}{RGB}{255,150,150}
\newcommand{\BL}[1]{{{\color{blue} #1}}}
\newcommand{\RD}[1]{{{\color{red} #1}}}
\newcommand{\REMOVEME}[1]{{{\color{black} #1}}}
\newcommand{\GR}[1]{{{\color{green} #1}}}
\newcommand{\YW}[1]{{{\color{yellow} #1}}}
\newcommand{\OR}[1]{{{\color{orange} #1}}}
\newcommand{\NF}[1]{\YW{{\bf Nick:} #1}}
\newcommand{\PvdH}[1]{{\OR{{\bf Pim:} #1}}}
\newcommand{\TM}[1]{\BL{{\bf Tobias:} #1}}
\newcommand{\MS}[1]{\GR{{\bf Markus: }#1}}
\newcommand{\ToDo}[1]{{\color{red} {\bf To do:} #1}}

\allowdisplaybreaks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%								Setup of the theorem environmnet								  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newtheorem{remark}{Remark}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{assumption}[definition]{Assumption}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%											Authors												  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Exact asymptotic expressions for local clustering in hyperbolic random graphs}
\author[1]{Nikolaos Fountoulakis}
\author[2]{Pim van der Hoorn}
\author[3]{Tobias M\"{u}ller}
\author[3]{Markus Schepers}

\affil[1]{University of Birmingham, School of Mathematics, United Kingdom}
\affil[2]{Northeastern University, Department of Physics, United States}
\affil[3]{University of Groningen, Bernoulli Institute for Mathematics, Computer Science and Artificial Inteligence, The Netherlands}

\begin{document}

\maketitle

\begin{abstract}
\PvdH{The abstract will be written at the end.}
\end{abstract}

\PvdH{Below is a list of high and low level to do items. Please add comments using the commands below.

\noindent Comment commands:\\
Markus: \textbackslash MS\\
Nikolaos: \textbackslash NF\\
Pim: \textbackslash PvdH\\
Tobias: \textbackslash TM

}

\ToDo{High level list:

\begin{itemize}
\item Streamline notations and writing of some sections
\item Finish introduction
\item Complete bibliography list
\item Add additional explanation for the Hyperbolic model.
\item Add additional observations based on our main result
\end{itemize}
}

\newpage

\tableofcontents

\newpage

\input{introduction}

\input{proof_overview}

\input{concentration_argument}

\input{clustering_infinite_model}

\input{transition}

\input{concentration}

\input{coupling}

\input{appendix}

\bibliographystyle{plain}
\bibliography{references}

\end{document}