### 2019-02-05 concentration_argument.tex

\section{Concentration of heights for vertices with degree $k$}\label{sec:concentration_argument}

In the proof of Proposition~\ref{prop:asymp} we used a result that allowed us to restrict integration over $y$ to the interval $[a(k)^-, a(k)^+]$, with
\[
	a(k)^\pm = 2\log\left(\frac{1 \pm C \sqrt{k \log(k)}}{\xi} \vee 1\right).
\]
The reason for this was that the integrand included the function $\rho(y,k) = \Prob{\Po(\mu(y)) = k}$, where $\Po(\lambda)$ denotes a Poisson random variable with expectation $\lambda$ and $\mu(y) = \mu(B_\Pcal(y)) = \xi e^{y/2}$, and Poisson random variables are well concentrated around their mean, i.e. around those values of $y$ for which $\mu(y) \approx k$.

In the remainder of this paper we will often encounter integrands involving, for some $\mu_n(y)$, the function $\Prob{\Po(\mu_n(y)) = k}$. In these case we want to be able to restrict our integration around those heights $y$ for which $\mu_n(y) \approx k_n$. We will refer to as \emph{concentration of heights arguments}. In this section we establish such results. We start with a concentration of heights lemma for the infinite model $\Ginf$ (Lemma~\ref{lem:concentration_argument}) and explain in Remark~\ref{rmk:concentration_argument} how such a result will be used throughout the paper. To obtain similar results for the other two models we have to first analyze the expected number of nodes in a typical neighborhood in these models. We do in Sections~\ref{ssec:average_degree_HP_n} and~\ref{ssec:average_degree_P_n}. We conclude this section with a general result that allows us to extend the concentration lemma for the infinite model to the hyperbolic random graph and finite box model.

\subsection{Concentration of heights argument for the infinite model}

%Recall that $\rho(y, k)$ is the probability density function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(B_\Pcal(y)) = \xie^{\frac{y}{2}}$. 
%We will consider two different types of sequences $\{k_n\}_{n \ge 1}$, those that are asymptotically bounded and those that diverge. We define
%\begin{equation}\label{eq:def_kappa_n}
%	\kappa_n := \begin{cases}
%		\log(n) &\mbox{if } k_n = \bigT{1},\\
%		\sqrt{k_n \log(k_n)} &\mbox{if } k_n = \omega(1).
%	\end{cases}
%\end{equation}
%\TM{ This is not a good definition. There are sequences that neither stay bounded nor tend to infinity, e.q.~odd values equal to 2 even values equal to $\log n$. How do you decide for a given $k_n$ which of the two cases you pick. To solve this we could either restrict to sequences that are either a) constant or b) tending to infinity; or we could make a cut-off at some specific, slow growing function such as $\log\log\log n$. }\PvdH{You are right, I overlooked this fact. I would opt to go with the setting where we consider only sequences that are either asymptotically bounded or tend to infinity. After the new proves are added we can see where we specify this.}
For any positive sequence $a_n \to \infty$ and $C > 0$ we define
\begin{equation}\label{eq:def_K_C_set}
	\Kcal_C(a_n) = \left\{y \in \R_+ : \frac{a_n - C \sqrt{a_n \log(a_n)}}{\xi} \vee 1 \le e^{\frac{y}{2}}
	\le \frac{a_n + C \sqrt{a_n \log(a_n)}}{\xi} \right\}.
\end{equation}
In addition we define
\begin{equation}\label{eq:def_K_C_n_set}
	\Kcal_{C,n}(a_n) := (-I_n, I_n] \times ((0,R_n] \cap \Kcal_{C}(a_n)),
\end{equation}
where $I_n := \frac{\pi}{2}e^{R_n/2}$. In addition, for a sequence $a_n \to \infty$, we shall often use the shorthand notation
\begin{equation}\label{eq:def_kappa_n}
	\kappa_n := \sqrt{a_n \log(a_n)}
\end{equation}

%Where we recall that $a \vee b = \max\{a,b\}$ and $a \wedge b = \min\{a,b\}$.

The next lemma states that for a large class of functions $h(y)$, to compute the integral 
\[
	\int_{0}^\infty \rho(y,k_n) h(y) e^{-\alpha y} \dd y
\]
%\TM{ This part is supposed to be for the infinite model, yet you integrate over the finite box $\Rcal_n$. Is this really meant?
%I would have expected we just integrate wrt.~$y$ and replace $f_{\alpha,\nu}(x,y)$ by $\alpha e^{-\alpha y}$, the exponential density.}\PvdH{I understand this confusion and agree that your suggestion makes more sense here. However, I choose this setup to minimize notations later. I am for changing the setup to what you are suggesting but I will have to spend some time first on checking how that will effect the definitions in later sections.}
it is enough to consider integration over $\Kcal_C(k_n)$ instead of $\R_+$. More precisely, for any $\ell_n = (1 + \smallO{1})k_n$ it is enough to consider $\Kcal_C(\ell_n)$. The flexibility of using $(1+\smallO{1})k_n$ instead of $k_n$ will prove useful in later sections.

\begin{lemma}\label{lem:concentration_argument}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\{k_n\}_{n \ge 1}$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$ and let $\ell_n = k_n(1 + \epsilon_n)$, with $\epsilon_n \to 0$. In addition let $\beta < \alpha$ and $h : \R_+ \rightarrow  \R$ be a any function such that $h(y) = g(y)e^{\beta y}$ with $|g(y)|$ uniformly bounded on $\R_+$. Then, if we define for $C > 0$
\[
	\lambda_n^\pm = (\ell_n \pm C \sqrt{\ell_n \log(\ell_n)}) \wedge \xi, 
	\quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right),
\] 
%\TM{ If $k$ is constant, then $\ell^- < 0$ and hence 
%$a^-$ is undefined! Need to change the definitions. }\PvdH{You are right. I changed it by taking the minimum with $\xi$.}
we have
\begin{equation}\label{eq:error_bound_int_rho_not_K}
	\int_{\R_+ \setminus \Kcal_{C}(\ell_n)} \rho(y,k_n) h(y) \alpha e^{-\alpha y} \dd y
	= \bigO{k_n^{-(1+C^2)/2}}
\end{equation}
%\TM{ As before, this is not a proper dichotomy. (unless we assume more about the sequences $k_n$, i.e.~either constant or tending to infinity }
as $n \to \infty$.

In particular, if $h_n(y)$ is a function such that $h_n(y) = \bigO{n^{-1} k_n^{s} e^{\beta y}}\rho(y,k_n)$ for some $s \in \R$ and $\beta < \alpha$ as $n \to \infty$. Then for $C > 0$ large enough we have
\[
	\lim_{n \to \infty} \int_{\Rcal_n \setminus \Kcal_{C,n}(\ell_n)} h_n(y) 
	f_{\alpha, \nu}(x,y) \dd x \dd y = 0,
\]
or equivalently,
\[
	\int_{\Rcal_n} \hspace{-5pt} h_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y
	= (1+\smallO{1}) \int_{\Kcal_{C,n}(\ell_n)} \hspace{-5pt} h_n(y) f_{\alpha, \nu}(x,y) \dd x \dd y.
\]
\end{lemma}

\begin{proof}

Note that in this case we write
\[
	\kappa_n = \sqrt{\ell_n \log(\ell_n)}.
\]
and recall (see proof of Proposition~\ref{prop:asymp}) that $\rho(y,k_n)$, as a function of $y$, is strictly increasing on $[0,a_n^-]$ and strictly decreasing on $[a_n^+,\infty)$. Therefore, by our assumption on $h(y)$,
\begin{align*}
	&\hspace{-20pt}\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho(y,k_n) 
		\alpha e^{-\alpha y} \dd y\\
    &= \bigO{1} \int_0^{a_n^-} e^{\beta y} \rho(y,k_n) \alpha e^{-\alpha y} \dd y 
    	+ \bigO{1}\int_{a_n^+}^{\infty} e^{\beta y} \rho(y,k_n) \alpha e^{-\alpha y}x \dd y \\
    &= \bigO{1} \int_0^{a_n^-} \rho(y,k_n) e^{-(\alpha-\beta) y} \dd y 
   		+ \bigO{1} \int_{a_n^+}^{\infty} \rho(y,k_n) e^{-(\alpha-\beta) y} \dd y\\
   	&\le \bigO{1}\rho(a_n^-,k_n)\int_0^{a_n^-} e^{-(\alpha-\beta) y} \, \dd y
   		+ \bigO{1} \rho(a_n^+,k_n) \int_{a_n^+}^{\infty} e^{-(\alpha-\beta) y} \, \dd y.
\end{align*}
Since $\alpha - \beta > 0$, we conclude that
\begin{equation}\label{eq:concentration_lemma_integral_bound}
	\int_{\R_+ \setminus [a_n^-, a_n^+]} h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y
	= \bigO{1} \left(\rho(a_n^-,k_n) + \rho(a_n^+,k_n)\right) 
\end{equation}

We shall now bound the terms $\rho(a_n^\pm,k_n)$, starting with $\rho(a_n^+,k_n)$.  Using Stirling's approximation $k! \sim \sqrt{2\pi} k^{k + 1/2} e^{-k}$ as $k \to \infty$ we write
\begin{align*}
	\rho(a_n^+,k_n) &= \frac{\mu(a_n^+)^{k_n}}{k_n!} e^{-\mu(a_n^+)} \\
	&\sim (2\pi)^{-1/2} k_n^{-1/2} \left(\frac{\mu(a_n^+)}{k_n}\right)^{k_n} e^{-(\mu(a_n^+) - k_n)}\\
	&= (2\pi)^{-1/2} k_n^{-1/2} 
		e^{-k_n\left(\frac{\mu(a_n^+)}{k_n} - 1 - \log\left(\frac{\mu(a_n^+)}{k_n}\right)\right)}.
\end{align*}

Since 
\[
	\frac{\mu(a_n^+)}{k_n} = \frac{\lambda_n^{+}}{k_n} = 1 + \epsilon_n + C \frac{\kappa_n}{k_n} 
	= 1 + \epsilon_n + C \sqrt{\frac{(1+\epsilon_n)\log((1+\epsilon_n)k_n)}{k_n}},
\]
and $x - \log(1 + x) \sim x^2/2$ as $x \to 0$, we get 
\begin{align*}
	\rho(a_n^+,k_n) 
	&\le \sqrt{2\pi} k_n^{-1/2} 
		e^{-k_n\left(\epsilon_n + C \frac{\kappa_n}{k_n} - \log\left(1 + \epsilon_n + C \frac{\kappa_n}{k_n}\right)\right)}\\
	&\sim (2\pi)^{-1/2} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n + C \kappa_n/k_n\right)^2}{2}} \\
	&= \bigO{k_n^{-(1+C^2)/2}},		\numberthis \label{eq:concentration_lemma_bound_an+}
\end{align*}
where for the last line we used that
\[
	-k_n \frac{\left(\epsilon_n + C \kappa_n/k_n\right)^2}{2} = -\frac{C^2}{2}\log(k_n) + \bigT{1}.
\]
A similar analysis as above yields
\begin{equation}\label{eq:concentration_lemma_bound_an-}
	\rho(a_n^-,k_n) \le \bigT{1} k_n^{-1/2} e^{-\frac{k_n \left(\epsilon_n - C \kappa_n/k_n\right)^2}{2}} = \bigO{k_n^{-(1+C^2)/2}}.
\end{equation}

Plugging~\eqref{eq:concentration_lemma_bound_an-} and~\eqref{eq:concentration_lemma_bound_an+}  into~\eqref{eq:concentration_lemma_integral_bound} yields the result. The second statement immediately follows from the first by choosing a large enough $C$ and observing that
\[
	\int_{\Rcal_n \setminus \Kcal_{C,n}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y
	\le n \int_{\R_+ \setminus [a_n^-, a_n^+]} h_n(y) \alpha e^{-\alpha y} \dd y.
\]
\end{proof}

\begin{remark}[Concentration of heights argument]\label{rmk:concentration_argument}
%\TM{ again, this is a loaded term. I would in the very least change to ``concentration of height'' everywhere }
Lemma~\ref{lem:concentration_argument} will prove very useful in the remainder of this paper since we often have to deal with integrands of the form $g_n(y) f_{\alpha,\nu}(x,y)$ where $g_n(y) = \bigO{n^{-1} k_n^{s}}e^{\beta y}\rho(y,k_n)$ for some $s \in \R$ and $\beta < \alpha$. In this case the lemma tells us that for a suitable $C > 0$ we only need to integrate over $\Kcal_C(k_n)$. In order words, we may always assume for $g_n(y)$ (for a penalty of $\smallO{1}$) that $k_n - C \kappa_n \le \xi e^{y/2} \le k_n + C \kappa_n$. We will refer to this as a \emph{concentration of heights argument}, e.g. \emph{by a concentration of heights argument}
\[
	\int_{\Rcal_n} g_n(y) \rho(y,k_n) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1+\smallO{1}) n \int_{\Kcal_C(k_n)} g_n(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y.
\]
\PvdH{ @All:It might be a good idea to add one more example of how the concentration argument is used in the paper. I someone finds a good example we use later on, please add it here.}
\end{remark}

We will later establish similar results for the cases where we consider the hyperbolic and finite box model. Here we have to consider Poisson distributions with mean $\mu(\BallHyp{y})$ and $\mu(\BallPon{y})$, respectively. For this the following, slightly more general version of Lemma~\ref{lem:concentration_argument} will be important.

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$, $k_n$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and let $\Kcal_{C,n}(\ell_n)$ be defined as in~\eqref{eq:def_K_C_n_set}. In addition, define $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, where $\hat{\mu}_n(y)$ satisfies, 
\[
	\hat{\mu}_n(y) = (1 + \phi_n(y))\mu(y),
\]
where $\phi_n(y)$ is a continuous differentiable function such that for some $0 < \varepsilon < 1$
\[
	\sup_{0 \le y \le (1 - \varepsilon)R_n} |\phi_n(y)| = 0.
\]
Then the following holds for some $C > 0$ large enough:
\begin{enumerate}
\item For any sequence of continuous functions $h_n(y) : \R_+ \to \R$ such that, uniformly on $(0,R_n]$, $h_n(y) = \bigO{k_n^s e^{\beta y} \hat{\rho}_n(y,k_n)}$, as $n \to \infty$, 
%\TM{ Why not adjust $h$ so that we incorporate $\hat\rho$ in line below. Then both items look more similar. }\PvdH{The reason for the bound on $h_n(y)$ in the first statement is that we want to use this for cases where $h_n(y) \le k_n^s \hat{\rho}_n(y,k_n)$ (see for instance the proof of Proposition~\ref{prop:convergence_average_clustering_P_n}.}  
for some $s < 2\alpha(2\alpha + 1)$ and $\beta < \alpha$, we have,
\[
	\int_{\Rcal_n} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Kcal_{C,n}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\item For any continuous bounded function $h(y)$,
\[
	\int_{\Kcal_{C,n}(\ell_n)} h(y) \hat{\rho}_n(y,k_n) f_{\alpha,\nu}(x,y) \dd x \dd y
	= (1+\smallO{1}) n \int_{0}^\infty h(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y,
\]
as $n \to \infty$.
\end{enumerate}
\end{lemma}

\begin{proof}
Similarly to the proof of Lemma~\ref{lem:concentration_argument} we define 
\[
	\lambda_n^\pm = (\ell_n \pm C \kappa_n) \wedge \xi, \quad \text{and} \quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi}\right).
\] 
%\TM{ Same problem as in L3.1. $\lambda^-$ can be negative, in which case $a^-$ is undefined.}

In addition we fix $\varepsilon^\ast > 0$ such that $\varepsilon < \min\{1- \frac{s}{2\alpha(2\alpha + 1)},\varepsilon, 1\}$ and. Then since $h_n(y) = \hat{\rho}_n(y,k_n) \bigO{k_n^s}$ we have
\begin{align*}
	k_n^s \int_{(1-\varepsilon^\ast)R_n}^{R_n} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	&= \bigO{1} \hat{\rho}_n((1-\varepsilon^\ast)R_n,k_n) k_n^s e^{-\alpha(1-\varepsilon^\ast)R_n} \\
	&= \bigO{\hat{\rho}_n((1-\varepsilon^\ast)R_n,k_n) k_n^s n^{-2\alpha(1 - \varepsilon^\ast)}} \\
	&= \smallO{ \hat{\rho}_n((1-\varepsilon^\ast)R_n,k_n) n^{\frac{s}{2\alpha + 1} - 2\alpha(1-\varepsilon^\ast)}} = \smallO{1},
\end{align*}
where the last step follows by our choice of $\varepsilon^\ast$. Hence it is enough to prove both statement on $\Rcal_n((1-\varepsilon)R_n, R_n)$ instead of $\Rcal_n$.
\PvdH{@All: This is where we use that $s < 2\alpha(2\alpha + 1)$. I think this can be removed by using a more careful bound but I am not sure this is needed.}

\paragraph{Proof of statement 1.}
Since we already showed that
\[
	\lim_{n \to \infty} k_n^s \int_{(1-\varepsilon)R_n}^{R_n} \hat{\rho}_n(y,k_n) e^{(\beta-\alpha) y} \dd y = 0,
\]
by our assumption on $h_n(y)$, it is enough to show that for sufficiently large $C > 0$,
\begin{equation}\label{eq:concentration_argument_an_-}
	\lim_{n \to \infty} k_n^s \int_0^{a_n^-} \hat{\rho}_n(y,k_n) e^{(\beta-\alpha) y} \dd y = 0,
\end{equation}
and
\begin{equation}\label{eq:concentration_argument_an_+}
	\lim_{n \to \infty} k_n^s \int_{a_n^+}^{(1-\varepsilon)R_n} \hat{\rho}_n(y,k_n) e^{(\beta-\alpha) y} \dd y = 0.
\end{equation}
%\TM{ Why can we change the upper limit of the integral? PLease elaborate (in the paper) }

For simplicity we write $\mu(y) := \mu_{\alpha,\nu}\left(\BallPo{y}\right)$.. Now fix some $0 < \delta < 1$ and let $n$ be large enough such that 
\begin{enumerate}
\item $\sup_{0 < y \le (1-\varepsilon)R_n} |\phi_n(y)| < \delta$,
\item $(\ell_n + C \kappa_n)(1 - \delta) > k_n$ and
\item $(\ell_n - C \kappa_n)(1 + \delta) < k_n$
\end{enumerate}

Next, recall that the function $\lambda \mapsto \Prob{\Po(\lambda) = k}$ is monotonic increasing on $[0,k]$ and monotonic decreasing on $[k, \infty)$. Then since for $n$ large enough we have 
\[
	\hat{\mu}_n(y) = \mu(y)(1 + \phi_n(y)) \ge \mu(y)(1 - \delta) \ge \mu(a_n^+)(1 - \delta)
	= (\ell_n + C \kappa_n)(1 - \delta) > k_n,
\]
it follows that 
\[
	\hat{\rho}_n(y,k_n) = \Prob{\Po(\hat{\mu}_n(y)) = k_n} \le \Prob{\Po(\mu(y)(1-\delta)) = k_n},
\]
for all $a_n^+ \le y \le (1-\varepsilon)R_n$. By making the change of variables $z = \mu^{-1}(\mu(y)(1-\delta)) = y + 2 \log(1-\delta)$ we then get
\begin{align*}
	k_n^s \int_{a_n^+}^{(1-\varepsilon)R_n} \hat{\rho}_n(y,k_n) e^{(\beta - \alpha)y} \dd y
	&\le k_n^s \int_{a_n^+}^{(1-\varepsilon)R_n} \Prob{\Po(\mu(y)(1-\delta)) = k_n} 
		e^{(\beta-\alpha) y} \dd y\\
	&= k_n^s (1-\delta)^{\beta - \alpha} \int_{a_n^+ + 2\log(1-\delta)}^{(1-\varepsilon)R_n + 2\log(1-\delta)} 
		\Prob{\Po(\mu(z)) = k_n} e^{(\beta - \alpha)z} \dd z\\
	&\le k_n^s (1-\delta)^{\beta - \alpha} \int_{a_n^+}^{\infty} \rho(z,k_n) e^{(\beta - \alpha)z} \dd z.
\end{align*}
Since Lemma~\ref{lem:concentration_argument} implies that for large enough $C$,
\[
	\lim_{n \to \infty} k_n^s \int_{a_n^+}^{\infty} \rho(z,k_n) e^{(\beta - \alpha)z} \dd z = 0,
\]
we have proven~\eqref{eq:concentration_argument_an_+}.

The proof of~\eqref{eq:concentration_argument_an_-} follows the same line of reasoning. This time we use that for $0 \le y \le a_n^-$,
\[
	\hat{\mu}_n(y) \le \mu(y)(1 + \delta) \ge \mu(a_n^-)(1 + \delta)
		= (\ell_n - C \kappa_n)(1 + \delta) < k_n,
\]
so that 
\[
	\hat{\rho}_n(y,k_n) = \Prob{\Po(\hat{\mu}_n(y)) = k_n} \le \Prob{\Po(\mu(y)(1+\delta)) = k_n}.
\]
Making a similar change of variables $z = y + 2 \log(1+\delta)$ we then get
\begin{align*}
	k_n^s \int_0^{a_n^-} \hat{\rho}_n(y,k_n) e^{(\beta - \alpha)y} \dd y
	&= k_n^s \int_0^{a_n^-}\Prob{\Po(\mu(y)(1-\delta)) = k_n} 
		e^{(\beta-\alpha) y} \dd y\\
	&\le k_n^s (1+\delta)^{\beta - \alpha} \int_0^{a_n^- + 2\log(1+\delta)} 
		\rho(z,k_n) e^{(\beta - \alpha)z} \dd z,
\end{align*}
and~\eqref{eq:concentration_argument_an_-} follows by another application of Lemma~\ref{lem:concentration_argument}.

\paragraph{Proof of statement 2.}

The proof of the second statement follows the same line of reasoning as above. First note that by Lemma~\ref{lem:concentration_argument} and the first statement, it is enough to show that
\[
	\int_{a_n^-}^{a_n^+} \hat{\rho}_n(y,k_n) h(y) e^{-\alpha y} \dd y
	= \left(1 + o(1)\right) \int_{a_n^-}^{a_n^+} \rho(z,k_n) h(z) e^{-\alpha z} \dd z.
\] 
or equivalently,
\[
	\int_{a_n^-}^{a_n^+} \rho(z,k_n) h(z) e^{-\alpha z} \dd z
	= \left(1 + o(1)\right)\int_{a_n^-}^{a_n^+} \hat{\rho}_n(y,k_n) h(y) e^{-\alpha y} \dd y.
\]

Using the change of variable $z = \mu^{-1}(\hat{\mu}_n(y))$ and writing $\hat{a}_n^\pm = \hat{\mu}_n^{-1}(\mu(a_n^\pm))$, we get
\begin{align*}
	&\hspace{-30pt}\int_{a_n^-}^{a_n^+} \rho(z,k_n) h(z) e^{-\alpha z} \dd z \\
	&= \int_{a_n^-}^{a_n^+} \Prob{\Po(\mu(z)) = k_n} h(z) e^{-\alpha z} \dd z\\
	&= \int_{\hat{a}_n^-}^{\hat{a}_n^+} \Prob{\Po(\hat{\mu}_n(y)) = k_n} h(\mu^{-1}(\hat{\mu}_n(y))) e^{-\alpha \mu^{-1}(\hat{\mu}_n(y))} 
		\frac{\hat{\mu}_n^\prime(y)}{\mu^\prime(\mu^{-1}(\hat{\mu}_n(y)))} \dd y,
\end{align*}
where the fraction is the last line follows from the chain rule and the fact that $(\mu^{-1})^\prime(t) = (\mu^\prime(\mu^{-1}(t)))^{-1}$.

Now recall that $\hat{\mu}_n(y) = \mu(y)(1 + \phi_n(y))$, with $\phi_n(y)$ continuous and differentiable and satisfying
$\max_{0 \le y \le (1 - \varepsilon)R_n} |\phi_n(y)| \to 0$. It follows that $\max_{0 \le y \le (1 - \varepsilon)R_n} |\phi_n^\prime(y)| \to 0$ which then implies that $\hat{\mu}_n^\prime(y) = (1 + \smallO{1})\mu^\prime(y)$ as $n \to \infty$, uniformly on $(0,(1-\varepsilon)R_n)$. In particular this holds uniformly on $[a_n^-, a_n^+]$. Next we note that $\mu^\prime(y) = \mu(y)/2$ from which it follows that uniformly on $[a_n^-, a_n^+]$,
\[
	\frac{\hat{\mu}_n^\prime(y)}{\mu^\prime(\mu^{-1}(\hat{\mu}_n(y)))}
	= \frac{2\hat{\mu}_n^\prime(y)}{\hat{\mu}_n(y)}
	= \frac{(1 + \smallO{1})2\mu^\prime(y)}{(1 + \smallO{1})\mu(y)}
	= (1+\smallO{1})
\]
In addition, uniformly on $[0,(1-\varepsilon)R_n]$,
\[
	\mu^{-1}(\hat{\mu}_n(y)) = 2 \log(\hat{\mu}(y)/\xi) = 2 \log(\mu(y)/\xi) + 2 \log(1 + \smallO{1}) = y + \smallO{1}.
\]
Therefore, by assumption on $h(y)$,
\begin{align*}
	&\hspace{-30pt}\int_{a_n^-}^{a_n^+} \rho(z,k_n) h(z) e^{-\alpha z} \dd z \\
	&= (1 + \smallO{1})e^{-\alpha b_n}\int_{\hat{a}_n^-}^{\hat{a}_n^+} \Prob{\Po(\hat{\mu}_n(y)) = k_n} h(y + \smallO{1})  
		e^{-\alpha y} \dd y\\
	&= (1 + \smallO{1}) \int_{\hat{a}_n^-}^{\hat{a}_n^+} \Prob{\Po(\hat{\mu}_n(y)) = k_n} h(y)  
		e^{-\alpha y} \dd y \\
	&= (1 + \smallO{1}) \int_{a_n^-}^{a_n^+} \Prob{\Po(\hat{\mu}_n(y)) = k_n} h(y)  
			e^{-\alpha y} \dd y,
\end{align*}
which finishes the proof.
\end{proof}

\subsection{Expected number of points in balls in the hyperbolic random graph}\label{ssec:average_degree_HP_n}

\TM{  the term ``average degree" does not fit very well with what goes on in this section. We are considering the expected number of points in a ball, but not yet computing the average degree of the graph. }\PvdH{I agree and have changed the header.}
Recall that under the coupling between the hyperbolic random graph and the finite box model, for two points $p, p^\prime$ with $y + y^\prime < R_n$, $p^\prime \in \BallHyp{p}$ exactly when $|x-x^\prime|_{\pi e^{R_n/2}} \le \Phi(r,r^\prime)$. In this setting, the coupling lemma (Lemma~\ref{lem:asymptotics_Omega_hyperbolic}) gives that  
\[
	e^{\frac{1}{2}(y+y^\prime)} - K e^{\frac{3}{2}(y+y^\prime) - R_n} \leq \Phi(r, r^\prime) 
		\leq  e^{\frac{1}{2}(y+y^\prime)} + K e^{\frac{3}{2}(y+y^\prime) - R_n},
\]
for some constant $K$. %\TM{ Say what $K$ is and what the restrictions on $r, r'$ are. } 
This result enables us to determine the measure of a ball around a given point $p=(0,y)$ which will be fairly useful in our subsequent analysis. Recall that the hyperbolic ball $\BallHyp{p}$ is a subset of $\Rcal_n$ and not of the hyperbolic disc $\Dcal_{R_n}$, i.e. the balls $\BallHyp{p}$ "live" in the finite box and not the hyperbolic disc.

\begin{lemma}\label{lem:average_degree_hyperbolic}
Let $\eps \in (0, 1)$. Then for all $0 \le y \le (1 - \eps)R_n$
\[
	 1 - \phi_{\H,n}^{(1)}(y) - \phi_{\H,n}^{(2)}(y) \le \frac{\Mu{\BallHyp{0,y}}}{\Mu{\BallPo{0,y}}} 
	 \le 1 - \phi_{\H,n}^{(1)}(y) + \phi_{\H,n}^{(2)}(y),
\]
where
\[
	\phi_{\H,n}^{(1)}(y) = \frac{2\alpha - 1 - 4\pi}{4\pi} e^{-(\alpha - \frac{1}{2})(R_n - y)} 
		+ \left(\alpha - \frac{1}{2}\right)\pi e^{-(\alpha - \frac{1}{2})R_n - y/2},
\]
and
\[
	\phi_{\H,n}^{(2)}(y) = 
	+ \begin{cases}
			\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
			&\mbox{if } 1/2 < \alpha < 3/2,\\
			\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
			\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
			&\mbox{if } \alpha > 3/2,
		\end{cases}
\]
with $K$ being the constant coming from the approximation of $\Phi$ in Lemma~\ref{lem:asymptotics_Omega_hyperbolic}.
%\TM{ We want to say in the lemma where $K$ comes from. (not everyone reads surrounding text, or reads linearly, in a paper as long as this one.) }
\end{lemma}

\begin{proof}
We perform the computation of $\Mu{\BallHyp{0,y}}$ by splitting the integration with respect to the height $y^\prime$ into the cases $y^\prime > R_n - y$ and $y^\prime \le R_n - y$, %\TM{ You have not said what $y^\prime$ is! } 
where for the latter we utilize Lemma~\ref{lem:asymptotics_Omega_hyperbolic}. Recall that $\Mu{\BallPo{0,y}} = \xi e^{y/2}$ where $\xi = \frac{4\alpha\nu}{\pi(2\alpha - 1)}$.

Note that $\BallHyp{(0,y)} \cap \Rcal_n ([R_n - y, R_n])= \Rcal_n([R_n-y,R_n])$. 
Thus, 
\begin{align*}
	&\hspace{-30pt}\Mu{\BallHyp{(0,y)} \cap \Rcal_n ([R_n - y, R_n])} \\
	&= \int_{R_n - y}^{R_n} \int_{I_n} f_{\alpha,\nu}(x^\prime, y^\prime) \dd x^\prime \dd y^\prime
		= \nu \alpha e^{R_n/2}\left(e^{-\alpha(R_n - y)} - e^{-\alpha R_n}\right)\\
	&= \Mu{\BallPo{0,y}} \frac{2\alpha - 1}{4\pi} \left( e^{-(\alpha - \frac{1}{2})(R_n - y)}
		- e^{-(\alpha - \frac{1}{2})R_n - y/2}\right) \numberthis \label{eq:mu_hyperbolic_ball_part_1}
\end{align*}

Next we will establish upper and lower bounds on $\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]}$. Using Lemma~\ref{lem:asymptotics_Omega_hyperbolic} we have 
\begin{align*}
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} 
	&\le \frac{2\nu \alpha}{\pi} \int_{0}^{R_n - y} \left(e^{\frac{y + y^\prime}{2}} + K e^{\frac{3}{2}(y + y^\prime) - R_n}\right)
		e^{-\alpha y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right) \\
	&\hspace{10pt}+ \frac{2\nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
\end{align*}
The last integral depends on the value of $\alpha$,
\[
	\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
	= \begin{cases}
		\frac{2}{3 - 2\alpha}\left(e^{(\frac{3}{2} - \alpha)(R_n - y)} - 1\right) &\mbox{if } 1/2 < \alpha < 3/2,\\
		R_n - y &\mbox{if } \alpha = 3/2,\\
		\frac{2}{2\alpha-3}\left(1 - e^{-(\alpha - \frac{3}{2})(R_n - y)}\right) &\mbox{if } \alpha > 3/2.
	\end{cases}
\]
Therefore we get
\begin{align*}
	&\hspace{-30pt}\frac{2\nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}
		\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\begin{cases}
		\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
		&\mbox{if } 1/2 < \alpha < 3/2,\\
		\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
		\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
		&\mbox{if } \alpha > 3/2.
	\end{cases}\\
	&= \Mu{\BallPo{0,y}} \phi_{\H,n}^{(2)}(y)
\end{align*}
and hence
\[
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} = \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)} 
	+ \phi_{\H,n}^{(2)}(y)\right).
\]
Combining this with~\eqref{eq:mu_hyperbolic_ball_part_1} yields the required upper bound.

The lower bound follows by observing that the only difference with the above computations is the change of sign in front of 
\[
	\frac{2 \nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime.
\]
\end{proof}

\subsection{Expected number of points in balls in the finite box model}\label{ssec:average_degree_P_n}

%\TM{ same comment Re: average degree. }
For the finite box model $\Gbox$ we obtain a similar result for expected size of the balls.

\begin{lemma}\label{lem:average_degree_P_n}
For all $p \in \Rcal_n$, such that $y > 2\log(\pi/2)$,
\[
	\mu_{\alpha,\nu}(\BallPon{p}) = \mu_{\alpha,\nu}(\BallPo{p})\left(1 - \phi_n(y)\right)
\]
%\TM{ where did the curly B's go? }\PvdH{Typo. Fixed}
where $\phi_n(y) \ge 0$ is given by
\[
	\phi_n(y) = \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)}e^{-(\alpha-\frac{1}{2})(R_n - y)}
	- \frac{(2\alpha - 1)\pi}{4\alpha}\left(\left(\frac{\pi}{2}\right)^{-2\alpha} 
	e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(\alpha - \frac{1}{2})R_n - \frac{y}{2}}\right).
\]
On the other hand, if $y \le 2 \log(\pi/2)$ then
\[
	\mu_{\alpha,\nu}(\BallPon{p}) = \mu_{\alpha,\nu}(\BallPo{p})\left(1 - e^{-(\alpha - \frac{1}{2})R_n}\right).
\]
\end{lemma}

\begin{proof}
First note that since we have identified the boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ we can assume, without loss of generality, that $p = (0,y)$. We then have that the boundaries of $\BallPon{p}$ are given by the equations $x^\prime = \pm e^{\frac{y+y^\prime}{2}}$, which intersect the left and right boundaries of $[-\frac{\pi}{2}e^{\frac{R_n}{2}}, \frac{\pi}{2}e^{\frac{R_n}{2}}]$ at height
\[
	h(y) = R_n + 2 \log\left(\frac{\pi}{2}\right) - y.
\]
Therefore, if $y \le 2 \log(\pi/2)$ this intersection occurs above the height $R_n$ of the box $\Rcal_n$ while in the other case the full region of the box above $h(y)$ is connected to $p$.

We will first consider the case where $y > 2 \log(\pi/2)$. Recall that $\mu_{\alpha,\nu}(\BallPo{p}) = \xi e^{\frac{y}{2}}$ where $\xi = \frac{4\alpha \nu}{(2\alpha - 1)\pi}$. Then, after some simple algebra, we have that
\begin{align*}
	\mu_{\alpha,\nu}(\BallPon{p})
	&= \int_0^{h(y)} \int_{-\frac{\pi}{2}e^{\frac{R_n}{2}}}^{\frac{\pi}{2}e^{\frac{R_n}{2}}} 
		\ind{|x^\prime| \le e^{\frac{y+y^\prime}{2}}} f_{\alpha,\nu}(x^\prime,y^\prime) \, dx^\prime \, dy^\prime\\
	&\hspace{10pt}+ \int_{h(y)}^{R_n} \int_{-\frac{\pi}{2}e^{\frac{R_n}{2}}}^{\frac{\pi}{2}e^{\frac{R_n}{2}}} 
		f_{\alpha,\nu}(x^\prime,y^\prime) \, dx^\prime \, dy^\prime\\
	&= \frac{2 \alpha \nu}{\pi} e^{\frac{y}{2}} \int_0^{h(y)} e^{-(\alpha - \frac{1}{2})y^\prime} \, dy^\prime
		+ \alpha \nu e^{\frac{R_n}{2}} \int_{h(y)}^{R_n} e^{-\alpha y^\prime} \, dy^\prime \\
	&= \xi e^{\frac{y}{2}}\left(1 - \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)} 
		e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)\\
	&\hspace{10pt}+ \nu e^{\frac{R_n}{2}}\left(\left(\frac{\pi}{2}\right)^{-2\alpha} e^{-\alpha(R_n - y)} 
		- e^{-\alpha R_n}\right)\\
	&= \mu_{\alpha,\nu}(\BallPo{p})\left(1 - \phi_n(y)\right).
\end{align*}
Since, for all $\alpha > \frac{1}{2}$,
\[
	\left(\frac{\pi}{2}\right)^{-(2\alpha - 1)} \ge \frac{(2\alpha - 1)\pi}{4\alpha} \left(\frac{\pi}{2}\right)^{-2\alpha}
\]
it follows that $\phi_n(y) \ge 0$.

When $y \le 2 \log(\pi/2)$ we have
\begin{align*}
	\mu_{\alpha,\nu}(\BallPon{p})
	&= \int_0^{R_n} \int_{-\frac{\pi}{2}e^{\frac{R_n}{2}}}^{\frac{\pi}{2}e^{\frac{R_n}{2}}} 
		\ind{|x^\prime| \le e^{\frac{y+y^\prime}{2}}} f_{\alpha,\nu}(x^\prime,y^\prime) \, dx^\prime \, dy^\prime\\
	&= \frac{2 \alpha \nu}{\pi} e^{\frac{y}{2}} \int_0^{R_n} e^{-(\alpha - \frac{1}{2})y^\prime} \, dy^\prime\\
	&= \mu_{\alpha,\nu}(\BallPo{p})\left(1 - e^{-(\alpha - \frac{1}{2})R_n}\right).
\end{align*}
\end{proof}

\subsection{Concentration of heights argument for hyperbolic and finite box model}

Let us define $\rho_{\Po}(y,k) = \Prob{\Po(\Mu{\BallHyp{y}}) = k}$ and $\rho_{\text{box}} = \Prob{\Po(\Mu{\BallPon{y}}) = k}$. Then since Lemma~\ref{lem:average_degree_P_n} implies that $\Mu{\BallHyp{(y)}}$ satisfies the requirements in Lemma~\ref{lem:concentration_argument_rho_approximation} while Lemma~\ref{lem:average_degree_hyperbolic} states that this holds for $\Mu{\BallPon{0,y}}$, we have the following important Corollary.

\begin{corollary}\label{cor:concentration_argument_other_models}
The statements of Lemma~\ref{lem:concentration_argument_rho_approximation} holds for the two distribution functions $\rho_{\Po}(y,k)$ and $\rho_{n}(y,k)$. 
\end{corollary}

%\begin{corollary}\label{cor:concentration_argument_other_models}
%Let $\hat{\rho}_n(y,k)$ be any of the two distribution functions $\rho_{\Po}(y,k)$ and $\rho_{n}(y,k)$ and let $\Kcal_{C,n}(\kappa_n)$ be defined as in~\eqref{eq:def_K_C_n_set}. Then, for any function $g_n(y)$ such that, uniformly on $(0, R_n]$, $g_n(y) = \bigO{k_n^{s} e^{\beta y}}$,
%%\TM{ Where did the $n^{-1}$ come from? Does not appear in L3.4. }\PvdH{I will elaborate on this once the statement of Lemma~\ref{lem:concentration_argument} is updated}, 
%as $n \to \infty$, for some $s < 2\alpha(2\alpha + 1)$ and $\beta < \alpha$,
%\[
%	\int_{\Rcal_n} g_n(y) \hat{\rho}_n(y,k_n) f_{\alpha,\nu}(x,y) \dd x \dd y
%	= (1+\smallO{1})n\int_{0}^\infty g_n(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y,
%\]
%for some $C > 0$ large enough.
%\end{corollary}

In particular we conclude that, similarly to the infinite limit model, concentration of heights arguments as described in Remark~\ref{rmk:concentration_argument} can be applied in the case of the hyperbolic random graphs and the finite box model.

\begin{remark}[Most frequent use of concentration of argument]
Note that it follows from Proposition~\ref{prop:asymp} that $P(y)$ satisfies the conditions in the both statements of Lemma~\ref{lem:concentration_argument_rho_approximation}. In particular, if $\hat{\rho}_n$ denote either $\rho_{\Po}$ or $\rho_{\text{box}}$, then as $n \to \infty$
\begin{equation}\label{eq:concentration_heights_argument_Py}
	\int_{\Rcal_n} P(y) \hat{\rho}_n(y,k_n) f_{\alpha,\nu}(x, y) \dd x \dd y
	= (1 + \smallO{1}) n \int_0^\infty P(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y.
\end{equation}
This is the form that will be use most frequently in the remainder of this paper.
\end{remark}