### 2020-03-12 proofs_finite_k.tex
Removed several LaTeX warnings

\section{Proofs of Theorem~\ref{thm:maincc} and Theorem~\ref{thm:mainkfixed}\label{sec:proofs_fixed_k}}

We will first derive Theorem~\ref{thm:mainkfixed}. It will turn out that Theorem~\ref{thm:maincc} has a quick derivation
assuming Theorem~\ref{thm:mainkfixed}.

\subsection{Clustering function for fixed \texorpdfstring{$k$}{k}, proving Theorem~\ref{thm:mainkfixed}}
%=======
%\subsection{Clustering function, proving Theorem~\ref{thm:mainkfixed}}
%In this subsection, we want to show that the clustering function of the KPKVB model $c(k; G(n;\alpha,\nu)) \xrightarrow{\Pee} \gamma(k)$ for any fixed $k$. The key idea is that for the poissonized KPKVB model the coupling with the box model is guaranteed to be exact (in the sense that it also preserves edges) for all vertices up to height $H = \frac{R}{4}$ (which is all we need to show the convergence of the clustering function for constant $k$).
%>>>>>>> Stashed changes

We will now show that the clustering function of the KPKVB model $c(k;G_n) \xrightarrow{\Pee} \gamma(k)$ for a fixed $k$. 
The key ideas are that the coupling of the Poissonized KPKVB model with the box model is guaranteed to be exact 
(in the sense that it also preserves edges) for all vertices up to height $R/4$; and that when computing 
the expected value clustering function $c(k;G)$ in the subgraph of the box model induced by all vertices up to height $R/4$
using the Campbell-Mecke formula we obtain integrals that are very similar to the expressions we found earlier for $\gamma(k)$.
%, the conditional expectation of 
%clustering coefficient of the typical point in the infinite limit model $\Ginf$ conditioned on the typical point having degree $k$.

%=======
%We will encounter the following models: we will start with the KPKVB model $G_n = G(n;\alpha,\nu)$, then the Poissonized KPKVB model $\GPo  = G_{Po}(n;\alpha,\nu)$, moving to the truncated Poissonized KPKVB model $\GPoH$ which is the subgraph of $\GPo$ induced on all vertices with height at most $H$ resp. radial coordinate at least $R-H$, then the coupling will bring us to the truncated box model $\GboxH$ which is the subgraph of the box model $G_{box}$ induced by all vertices with height at most $H$.
%>>>>>>> Stashed changes

We will repeatedly rely on the following observation.

\begin{lemma}\label{lem:ckGckH}
Let $k\geq 2$ and let $G, H$ be graphs such that $G$ is an induced subgraph of $H$, or vice versa.
Then

$$ \left| c(k;G) - c(k;H) \right| \leq \frac{6|E(G)\Delta E(H)|}{N_G(k) - 2|E(G)\Delta E(H)|}, $$

\noindent
provided $N_G(k) > 2|E(G)\Delta E(H)|$.
\end{lemma}

\begin{proof} 
We observe that

$$ \begin{array}{rcl} 
|c(k; G)-c(k;H)| 
& = & \displaystyle
\left| \sum_{\substack{v\in V(G), \\ \text{deg}_G(v)=k}} \frac{c_G(v)}{N_G(k)} - 
\sum_{\substack{v\in V(H), \\ \text{deg}_{H}(v)=k}} \frac{c_{H}(v)}{N_{H}(k)} \right| \\[8ex]
& = & \displaystyle 
\left| \frac{1}{N_G(k)} \left( \sum_{\substack{v\in V(G)\setminus V(H), \\ \text{deg}_G(v)=k} } c_G(v)
+ \sum_{\substack{v\in V(G)\cap V(H), \\ \text{deg}_G(v)=k, \\ \text{deg}_{H}(v)\neq k}} c_G(v) \right) \right. \\
& & \displaystyle
- \frac{1}{N_{H}(k)} \left(\sum_{\substack{ v\in V(H)\setminus V(G), \\ \text{deg}_{H}(v)=k} } c_{H}(v)
+ \sum_{\substack{v\in V(G)\cap V(H), \\ \text{deg}_G(v)\neq k, \\ \text{deg}_{H}(v)=k}} c_{H}(v) \right) \\
& & \displaystyle  
\left. + \left(\frac{1}{N_G(k)}-\frac{1}{N_{H}(k)} \right) \cdot 
\left( \sum_{\substack{v \in V(G)\cap V(H), \\ \text{deg}_G(v)=\text{deg}_{H}(v)=k}}
c_G(v) \right) \right| \\[8ex]
 & \leq & \displaystyle
 \frac{2|E(G)\Delta E(H)|}{N_G(k)} + \frac{2|E(G)\Delta E(H)|}{N_{H}(k)} + 
 \frac{|N_{H}(k) - N_G(k)|}{N_G(k)\cdot N_{H}(k)} \cdot N_{H}(k) \\[5ex]
 & = & \displaystyle 
 \frac{2|E(G)\Delta E(H)|}{N_G(k)} + \frac{2|E(G)\Delta E(H)|}{N_{H}(k)} + \frac{|N_{H}(k) - N_G(k)|}{N_G(k)} \\[5ex]
 & \leq & \displaystyle
 \frac{2|E(G)\Delta E(H)|}{N_G(k)} + \frac{2|E(G)\Delta E(H)|}{N_{H}(k)} + \frac{2|E(G)\Delta E(H)|}{N_G(k)} \\[5ex]
 & \leq & \displaystyle
\frac{6 |E(G)\Delta E(H)|}{N_G(k)-2|E(G)\Delta E(H)|}.
\end{array} $$

\noindent
(In the second line we use that $\text{deg}_G(v) = \text{deg}_{H}(v)$ in fact 
implies that $c_G(v) = c_{H}(v)$ since one of $G,H$ is an induced subgraph of the other. 
In the third line we use that clustering coefficients $c_G(v), c_H(v)$ take values in $[0,1]$, and if either 
$\text{deg}_{G}(v) \neq \text{deg}_{H}(v)$ or $v \in V(G)\Delta V(H)$ and $v$ has degree $K$ in whichever of $G, H$ it belongs to 
then at least one edge of $E(G)\Delta E(H)$ is incident with $v$, and that every edge in 
$E(G)\Delta E(H)$ only affects the status of its two incident vertices.
For the fifth line we used that $|N_G(k)-N_H(k)| \leq |\{ v\in V(G) : \text{deg}_G(v) = k \} \Delta 
\{ v\in V(H) : \text{deg}_H(v) = k \}| \leq 2|E(G)\Delta E(H)|$ for similar reasons.
In the last line we used that $N_H(k) \geq N_G(k) - |N_G(k)-N_H(k)| \geq N_G(k) - 2|E(G)\Delta E(H)|$.)
\end{proof}

\begin{lemma}\label{lem:NDGnGPo}
%Under the standard coupling of $G_n$ and $\GPo$, we have 
$|E(G_n) \Delta E(\GPo) | = o(n)$ a.a.s. 
\end{lemma}

\begin{proof}
Let us fix some $\eps>0$ and write

$$G_- := G((1-\eps)n, \alpha, (1-\eps)\nu ), \quad G_+ := G((1+\eps)n, \alpha, (1+\eps)\nu ). $$

\noindent
(We ignore rounding issues, i.e.~the issue that $(1-\eps)n, (1+\eps)n$ may not be integers, 
to avoid notational burden. We leave the straightforward details of adapting our arguments below to deal with it to the reader.)

Observe that the vertices of $G_-, G_+, G_n, \GPo$ all live on the same hyperbolic disk, of radius 
$R = 2\ln(n/\nu)$. 
We consider the standard coupling where we have an infinite supply of i.i.d.~points $u_1, u_2, \dots$ chosen according
to the $(\alpha,R)$-quasi uniform distribution, the vertices of $G_n = G(n;\alpha,\nu)$ are $u_1,\dots, u_n$, 
the vertices of $G_-$ are $u_1,\dots, u_{(1-\eps)n}$, the vertices of $G_+$ are $u_1,\dots, u_{(1+\eps)n}$ and 
the vertices of $\GPo$ are $u_1,\dots, u_N$ with $N\isd \Po(n)$ independently of
$u_1, u_2, \dots$.

As $N\isd\Po(n)$, by Chebychev's inequality, we have that $|N-n| < \eps n$ a.a.s.
So in particular, under our coupling we have $G_- \subseteq G_n, \GPo \subseteq G_+$ a.a.s. 
We now point out that, by the results of Gugelmann et al.~on the average degree (\cite{gugelmann2012random}, Theorem 2.3), 
we have

$$|E(G_-)| = (1-\eps)^2\cdot \frac{4\nu\alpha^2}{\pi(2\alpha-1)^2} \cdot n + o(n), \quad 
|E(G_+)| = (1+\eps)^2 \cdot \frac{4\nu\alpha^2}{\pi(2\alpha-1)^2} \cdot n + o(n) \quad \text{ a.a.s. } $$

%$G_-$ hand $G_+$ both have $4\nu\alpha^2/\pi(2\alpha-1)^2 \cdot n + o(n)$ many edges (a.a.s.).
It follows that

$$ | E(G_n) \Delta E(\GPo) | \leq |E(G_+)\setminus E(G_-)|
= \eps \cdot \frac{16\nu\alpha^2}{\pi(2\alpha-1)^2}\cdot n + o(n) \text{ a.a.s. }
$$

This holds for every fixed $\eps>0$. Sending $\eps\searrow 0$, concludes the proof of the lemma.
\end{proof}

Next, let us recall that by the results of Gugelmann et al.~on the degree sequence
(\cite{gugelmann2012random}, Theorem 2.2) we have that

\begin{equation}\label{eq:gugeldegseqaap} 
\frac{N_{G_n}(k)}{n} \xrightarrow[n\to\infty]{\Pee} \pmf(k), 
\end{equation}

\noindent
for every fixed $k$.
In particular $N_{G_n}(k) = \Omega(n)$ a.a.s. Combining this with lemmas~\ref{lem:ckGckH} and~\ref{lem:NDGnGPo} we obtain:

\begin{corollary}\label{cor:pok} 
%Under the standard coupling of $G_n$ and $\GPo$, we have
For every fixed $k\geq 2$, we have

$$ c(k; G_n)-c(k; \GPo) \xrightarrow[n\to\infty]{\Pee} 0, \quad\text{ and }\quad
\frac{N_{\GPo}(k)}{n} \xrightarrow[n\to\infty]{\Pee} \pmf(k). $$

\end{corollary}

\noindent
(For the second statement we use that $N_{G_n}(k) - 2|E(G_n)\Delta E(\GPo)| \leq N_{\GPo}(k) \leq N_{G_n}(k) + 2|E(G_n)\Delta E(\GPo)|$.)

In the remainder of this section, we'll denote by $\GboxH$ the subgraph of $\Gbox$ induced by all 
vertices $(x,y) \in \Vbox = \Pcal \cap \Rcal$ of height at most $R/4$.

\begin{lemma}
Under the coupling provided by Lemma~\ref{lem:coupling_hyperbolic_poisson}, a.a.s., 
$\GboxH$ is an induced subgraph of $\GPo$ and $|E(\GPo) \setminus E(\GboxH)| = o(n)$.
\end{lemma}

\begin{proof}
We remind the reader that under the coupling of Lemma~\ref{lem:coupling_hyperbolic_poisson}, 
we can view $\Gbox$ and $\GPo$ as having the same vertex set $\Vbox = \Pcal \cap \Rcal$; and 
two points $p = (x,y), p' = (x',y') \in \Vbox$ are joined by an edge in $\Gbox$ if 
$|x-x'|_{\pi e^{R/2}} \leq e^{(y+y')/2}$, while $p,p'$ are joined by an edge in $\GPo$ 
if either $y+y'\geq R$ or $y+y'<R$ and $|x-x'|_{\pi e^{R/2}} \leq \Phi(y,y')$ with $\Phi$ as provided by~\eqref{eq:def_Omega_hyperbolic}.
It follows immediately from Lemma~\ref{lem:coupling_edges} that $\GboxH$ is an induced subgraph of $\GPo$, a.a.s., as claimed.

Fix $\eps > 0$, and let $X$ denote the number points of $\Vbox$ with $y$-coordinate $\geq (1-\eps) R$. 
Then $X$ is a Poisson random variable with mean

$$ \begin{array}{rcl} 
\Ee X 
& = & \displaystyle \mu\left( (-\frac{\pi}{2} e^{R/2},\frac{\pi}{2} e^{R/2}] \times [(1-\eps) R,R] \right)
= \int_{-\frac{\pi}{2} e^{R/2}}^{\frac{\pi}{2} e^{R/2}}\int_{(1-\eps)R}^R \left(\frac{\nu\alpha}{\pi}\right) e^{-\alpha y}\;dy\;dx \\[4ex]
& = & O( e^{R/2-(1-\eps)\alpha R} ) = o(1), 
\end{array} $$

\noindent
the last equality holding provided $\eps$ was chosen sufficiently small (using that $\alpha > 1/2$).
We conclude that, a.a.s., there are no vertices of height $\geq (1-\eps) R$.

Let $Y$ denote the number of pairs of vertices $p = (x,y), p'=(x',y') \in \Vbox$ with $y+y' \geq R$.
Then, by the Campbell-Mecke formula

$$ \begin{array}{rcl} \Ee Y 
& = & \displaystyle
\int_\Rcal\int_\Rcal 1_{\{y+y'\geq R\}} \mu(dp')\mu(dp) \\[4ex]
& = & \displaystyle
\int_{-\frac{\pi}{2}e^{R/2}}^{\frac{\pi}{2}e^{R/2}}\int_0^R
\int_{-\frac{\pi}{2}e^{R/2}}^{\frac{\pi}{2}e^{R/2}}\int_{R-y}^R \left(\frac{\nu\alpha}{\pi}\right)^2 e^{-\alpha(y+y')}dy'dx'dydx \\[4ex]
%= n^2 \int_0^R \int_{R-y}^R \alpha^2 e^{-\alpha(y+y')}dy'dy
& = & O( R e^{(1-\alpha)R} ) = o( n ),
\end{array} $$

\noindent
the last equality holding because $\alpha > 1/2$ and $n = \nu e^{R/2}$.
In particular, by Markov's inequality, $Y = o(n)$ a.a.s.

Now let $Z$ denote the number of pairs of vertices $p = (x,y), p'=(x',y')$
for which $y+y'<R, y < (1-\eps)R, R/4 \leq y' < (1-\eps)R$ and $|x-x'|_{\pi e^{R/2}} < \Phi(y,y')$.
By Lemma~\ref{lem:asymptotics_Omega_hyperbolic} we have that $\Phi(y,y') = O( e^{(y+y')/2} )$ 
for all such $y,y'$.
By Campbell-Mecke we can write

$$ \begin{array}{rcl} 
\Ee Z 
& = & \displaystyle
\int_{-\frac{\pi}{2}e^{R/2}}^{\frac{\pi}{2}e^{R/2}}\int_0^{(1-\eps)R}
\int_{-\frac{\pi}{2}e^{R/2}}^{\frac{\pi}{2}e^{R/2}}\int_{R/4}^{(1-\eps)R} 
1_{\{|x-x'|_{\pi e^{R/2}}<\Phi(y,y'), y+y'<R\}}\left(\frac{\nu\alpha}{\pi}\right)^2 e^{-\alpha(y+y')}dy'dxdydx' \\[4ex]
& = & \displaystyle
O\left( e^{R/2} \int_0^{(1-\eps)R} \int_{R/4}^{(1-\eps)R} e^{(1/2-\alpha)(y+y')} dy'dy \right) \\[4ex]
& = & \displaystyle
O\left( e^{R/2 + (1/2-\alpha) R/4} \right) = o(n).
\end{array} $$

\noindent
Hence also $Z = o(n)$ a.a.s.

This concludes the proof as we've now shown that under the stated coupling, a.a.s., $\GboxH$ and $\GPo$ differ by only $o(n)$ edges.
\end{proof}

Analogously to Corollary~\ref{cor:pok} we obtain:

\begin{corollary}\label{cor:GPoGboxH}
For every fixed $k\geq 2$ we have

$$c(k;\GPo) - c(k;\GboxH) \xrightarrow[n \rightarrow \infty]{\Pee} 0, \quad\text{ and }\quad  
\frac{N_{\GboxH}(k)}{n} \xrightarrow[n\to\infty]{\Pee} \pmf(k). $$

\end{corollary}

\begin{lemma}\label{lem:ckGboxH} 
For every fixed $k\geq 2$ we have

 $$ c(k;\GboxH) \xrightarrow[n \rightarrow \infty]{\Pee}  \gamma(k). $$

 \end{lemma}

 \begin{proof}
 We write $\Rcal_- := \Rcal \cap (\eR\times[0,R/4]) = (-\frac{\pi}{2}e^{R/2}, \frac{\pi}{2} e^{R/2}] \times [0,R/4]$ and set

 $$ 
 X := \sum_{v \in N_{\GboxH}(k)} c(v) = \sum_{v \in \Pcal \cap \Rcal_-} c_{\GboxH}(v) \cdot 1_{\{\text{deg}_{\GboxH}(v)=k\}}.
 $$

 \noindent
 By the Campbell-Mecke formula

 $$ \Ee X = \int_{\Rcal_-} 
 \Ee_\Pcal\left[ c_{\GboxH^z}(z) \cdot 1_{\{\text{deg}_{\GboxH^z}(z)=k\}}\right] \mu(dz), $$

 \noindent
 where $\GboxH^z$ denotes the graph we get by adding $z$ as an additional vertex to $\GboxH$, and adding edges between $z$ and the other 
 vertices as per the connection rule (for $\Gbox$).
 Spelling out the intensity measure $\mu$, plus symmetry considerations, gives

 $$ \begin{array}{rcl}
     \Ee X 
     & = & \displaystyle 
     \int_{-\frac{\pi}{2}e^{R/2}}^{\frac{\pi}{2}e^{R/2}} \int_0^{R/4}
     \Ee_\Pcal\left[ c_{\GboxH^{(x,y)}}((x,y)) \cdot 1_{\{\text{deg}_{\GboxH^{(x,y)}}((x,y))=k\}}\right] 
     \left(\frac{\nu\alpha}{\pi}\right) e^{-\alpha y}\;dy\;dx \\[6ex]
     & = & \displaystyle
     n \dot \int_0^{R/4}
     \Ee_\Pcal\left[ c_{\GboxH^{(0,y)}}((0,y)) \cdot 1_{\{\text{deg}_{\GboxH^{(0,y)}}((0,y))=k\}}\right] 
     \alpha e^{-\alpha y}\;d y \\[6ex]
     & = & \displaystyle
     n \cdot \int_0^{R/4} \Ee_\Pcal\left[ c_{\GboxH^{(0,y_0)}}((0,y_0)) {\Big|} \text{deg}_{\GboxH^{(0,y_0)}}((0,y_0))=k \right] \cdot \\[6ex]
     & & \displaystyle \hspace{6ex} 
     %\cdot 
     \Pee\left[ \text{deg}_{\GboxH^{(0,y_0)}}((0,y_0))=k  \right] \alpha e^{-\alpha y_0} \;dy_0.
    \end{array}
$$

\noindent
The random variable $\text{deg}_{\GboxH^{(0,y_0)}}((0,y_0))$ follows a Poisson distribution with mean

$$ \begin{array}{rcl} 
\Ee\left[ \text{deg}_{\GboxH^{(0,y_0)}}((0,y_0))\right] 
& = & \displaystyle  
\mu(\BallPo{(0,y_0)}\cap\Rcal_-) 
= \int_0^{R/4} \int_{-e^{(y+y_0)/2}}^{e^{(y+y_0)/2}} \left(\frac{\nu\alpha}{\pi}\right) e^{-\alpha y}\;dx\;dy \\
& = & \xi e^{y_0/2} \cdot (1-e^{(1/2-\alpha)R/4}). 
\end{array} $$

\noindent
Hence, for every fixed $y_0$ and $k$, we have that

$$ \Pee\left[ \text{deg}_{\GboxH^{(0,y_0)}}((0,y_0))=k\right] %= \Pee( \Po( \mu(\BallH{(0,y_0)}) = k ) 
\xrightarrow[n\to\infty]{} 
%\Pee( \Po( \mu(\BallPo{(0,y_0)})) = k ) = 
\Pee( \Po(\xi e^{y_0/2})=k ) = \rho(y_0,k).$$

%\noindent
%with $D$ denoting the (unrestricted) degree of the typical point in the limit model $\Ginf$.

Next we remark that, analogously to the argument given in the beginning of Section~\ref{sec:42}, we have

$$ %\Ee[ c_{\GboxH^{(0,y_0)}}((0,y_0)) | \text{deg}_{\GboxH^{(0,y_0)}}((0,y_0))=k ] 
\Ee\left[ c_{\GboxH^{(0,y_0)}}((0,y_0)) {\Big|} \text{deg}_{\GboxH^{(0,y_0)}}((0,y_0))=k \right] 
= \Pee( w_1 \in \BallPo{w_2} ) =: P_n(y_0), $$

\noindent
with $w_1 = (x_1, y_1), w_2 = (x_2, y_2)$ chosen independently from $\BallPo{(0,y_0)} \cap \Rcal_-$ according to the probability 
measure we get by renormalizing $\mu$, i.e.~with pdf $f_\mu \cdot 1_{\BallPo{(0,y_0)} \cap \Rcal_-} / \mu(\BallPo{(0,y_0)}\cap \Rcal_-)$.
By considerations completely analogous to those following Lemma~\ref{lem:Paneq1}, the random variables $y_1, y_2$ both follow
a truncated exponential distribution with parameter $\alpha - 1/2$ truncated at height $R/4$
(i.e.~with density $1_{\{y_i \leq R/4\}} \cdot (\alpha-1/2) e^{(1/2-\alpha)y_i} / (1-e^{(1/2-\alpha)R/4})$) and, given the 
values of $y_0, y_1, y_2$, each $x_i$ is chosen uniformly on the interval $[-e^{(y_0+y_i)/2}, e^{(y_0,y_i)/2}]$.
In particular

$$ P_n(y_0) = 
\left(\frac{\alpha-1/2}{1-e^{(1/2-\alpha)R/4}}\right)^2 \int_0^{R/4}\int_0^{R/4} P(y_0, y_1, y_2) e^{(1/2-\alpha)(y_1+y_2)}\;dy_1\;dy_2,
$$

\noindent
with $P(.,.,y.)$ as defined in the paragraph following Lemma~\ref{lem:Paneq1}.
(That is, $P(y_0, y_1, y_2)$ is the probability that $|x_1-x_2| < e^{(y_1+y_2)/2}$, where $x_1, x_2$ are independent with $x_i$ uniform 
on the interval $[-e^{(y_0+y_i)/2}, e^{(y_0+y_i)/2}]$). 
It follows that, for any fixed $y_0$, we have

$$ P_n(y_0) \xrightarrow[n\to\infty]{} 
(\alpha-1/2)^2\int_0^\infty\int_0^\infty P(y_0, y_1, y_2) e^{(1/2-\alpha)(y_1+y_2)}\;dy_1\;dy_2 = P(y_0). $$

\noindent
(Applying monotone convergence to justify the convergence of the integral as $n \to \infty$.)

Since (expected) clustering coefficients and probabilities are between zero and one and $\alpha e^{\alpha y_0}$ is integrable, we can now apply the dominated 
convergence theorem to obtain that

\begin{equation}\label{eq:EXn} 
\frac{\Ee X}{n} \xrightarrow[n\to\infty]{} \int_0^\infty P(y_0) \rho(y_0,k)\alpha e^{-\alpha y_0}\;dy_0 = \pmf(k) \cdot \gamma(k).  
\end{equation}

\noindent
(Applying~\eqref{eq:gammakint} for the last equality.)

Next, we turn attention to $X(X-1) = \sum_{u\neq v \in N_{\GboxH}(k)} c(v)c(u)$.
Another application of Campbell-Mecke shows that

$$
\Ee X(X-1) 
= 
\int_{\Rcal_-}\int_{\Rcal_-} \Ee\left[ c_{\GboxH^{z,z'}}(z)c_{\GboxH^{z,z'}}(z') \cdot 
1_{\{\text{deg}_{\GboxH^{z,z'}}(z)=\text{deg}_{\GboxH^{z,z'}}(z')=k\}}\right]\mu(dz)\mu(dz'),
$$

\noindent
with $\GboxH^{z,z'}$ denoting the graph we get by adding $z, z'$ as additional vertices to $\GboxH$.
Now note that if $z=(x,y)$ and $z'=(x',y')$ satisfy $|x-x'|_{\pi e^{R/2}} > 2 e^{R/4}$ then 
the neighbourhoods of $z,z'$ are determined by the points of the Poisson process $\Pcal$ in disjoint areas
of the plane.
This implies that, provided $|x-x'|_{\pi e^{R/2}} > 2 e^{R/4}$:

\begin{equation}\label{eq:noot1} 
\begin{array}{c} 
\Ee\left[ c_{\GboxH^{z,z'}}(z)c_{\GboxH^{z,z'}}(z') \cdot 
1_{\{\text{deg}_{\GboxH^{z,z'}}(z)=\text{deg}_{\GboxH^{z,z'}}(z')=k\}}\right] \\
= \\
\Ee\left[ c_{\GboxH^{z}}(z) \cdot 
1_{\{\text{deg}_{\GboxH^{z}}(z)=k\}}\right] \cdot 
\Ee\left[ c_{\GboxH^{z'}}(z') \cdot 
1_{\{\text{deg}_{\GboxH^{z'}}(z')=k\}}\right].
\end{array} 
\end{equation}

\noindent
On the other hand, the LHS of~\eqref{eq:noot1} is always between zero and one, also if $|x-x'|_{\pi e^{R/2}} \leq 2 e^{R/4}$. 
We may conclude that

$$ \begin{array}{rcl} 
\Ee X(X-1) 
& \leq & \displaystyle  
\int_{\Rcal_-}\int_{\Rcal_-} \Ee\left[ c_{\GboxH^{z}}(z) \cdot 
1_{\{\text{deg}_{\GboxH^{z}}(z)=k\}}\right] %\\
%& & \displaystyle 
\cdot 
\Ee\left[ c_{\GboxH^{z'}}(z') \cdot 
1_{\{\text{deg}_{\GboxH^{z'}}(z')=k\}}\right] \mu(dz)\mu(dz') \\[8ex]
& & \displaystyle 
+ \int_{\Rcal_-}\int_{\Rcal_-} 1_{\{|x-x'|\leq 2e^{R/4}\}} \mu(dz)\mu(dz') \\[8ex]
& = & \displaystyle
\left( \int_{\Rcal_-} \Ee\left[ c_{\GboxH^z}(z) \cdot 1_{\{\text{deg}_{\GboxH^z}(z)=k\}}\right] \mu(dz) \right)^2
+ O( e^{3R/4} ) \\[8ex] 
& = & \displaystyle
\left(\Ee X\right)^2 + O( n^{3/2} ).
\end{array} $$

\noindent
Combining this with~\eqref{eq:EXn}, it follows that 
$\Var X = \Ee X(X-1) + \Ee X - \left(\Ee X\right)^2 = o\left( \left(\Ee X\right)^2 \right)$.
 By Chebychev's inequality, we therefore have

 $$ X = n \cdot \gamma(k) \cdot \pmf(k) + o(n) \text{ a.a.s. } $$

In combination with Corollary~\ref{cor:GPoGboxH} (second limit) we can conclude that

$$ c(k;\GboxH) = \frac{X}{N_{\GboxH(k)}} \xrightarrow[n\to\infty]{\Pee} \gamma(k), $$

\noindent
as desired.
 \end{proof}

\begin{proofof}{Theorem~\ref{thm:mainkfixed}}
For completeness, we point out that Theorem~\ref{thm:mainkfixed} follows immediately from 
Corollaries~\ref{cor:pok},~\ref{cor:GPoGboxH} and Lemma~\ref{lem:ckGboxH}.
\end{proofof}

\subsection{Overall clustering coefficient, proving Theorem~\ref{thm:maincc}}

% 
% 
% 
% In this subsection, we will infer that the clustering coefficient of the KPKVB random graph 
% $c(G_n) \xrightarrow{\Pee} \gamma$ by using that $c(k;G_n) \xrightarrow{\Pee} \gamma(k)$ and $N_{G_n}(k)/n \xrightarrow{\Pee} \pmf(k)$ 
% for all fixed $k$.

\begin{proofof}{Theorem~\ref{thm:maincc}}
Recall in Section~\ref{sec:Ginf},  we {\em defined} $\pmf(k) := \Pee(D=k), \gamma := \Ee C, \gamma(k) := \Ee(C|D=k)$ with $D$ the degree
and $C$ the clustering coefficient of the ``typical point'' in the infinite limit model $\Ginf$. We can write

$$ \gamma = \Ee C = \sum_{k\geq 2} \Ee\left( C | D=k \right) \Pee( D=k ) = \sum_{k\geq 2} \gamma(k) \cdot \pmf(k). $$

For the KPKVB random graph, or any graph for that matter, we have the similar relation

$$ c(G_n) = \sum_{k\geq 2} c(k;G_n) \cdot (N_{G_n}(k)/n). $$

By Theorem~\ref{thm:mainkfixed} and~\eqref{eq:gugeldegseqaap} we have, for any fixed $k\geq 2$:

\begin{equation}\label{eq:slutskyUB} 
 c(G_n) \geq \sum_{k=2}^K c(k;G_n) \cdot (N_{G_n}(k)/n) \xrightarrow[n\to\infty]{\Pee} \sum_{k=2}^K \gamma(k)\cdot \pmf(k), 
\end{equation}

\noindent
where Slutsky's theorem justifies the convergence in probability.
On the other hand we have

\begin{equation}\label{eq:slutskyLB} 
c(G_n) \leq \sum_{k=2}^K c(k;G_n) \cdot (N_{G_n}(k)/n) + \sum_{k>K} (N_{G_n}(k)/n)
\xrightarrow[n\to\infty]{\Pee} \sum_{k=2}^K \gamma(k)\cdot \pmf(k) + \sum_{k>K} \pmf(k), 
\end{equation}

\noindent
where the convergence in probability can be justified using Slutsky's theorem together 
with the fact that $\sum_{k=0}^\infty \pmf(k) = 1$ (one convenient way to convince oneself that this is true, is to note that 
$D$, the degree of the typical point, is a.s.~finite). In more detail,

$$\sum_{k>K} (N_{G_n}(k)/n) 
= 1 - \sum_{k=0}^K (N_{G_n}(k)/n) \xrightarrow[n\to\infty]{\Pee} 1 - \sum_{k=0}^K \pmf(k) = \sum_{k>K} \pmf(k). $$

The result follows from~\eqref{eq:slutskyLB} and~\eqref{eq:slutskyUB}, by sending $K\to\infty$. 
\end{proofof}