### 2020-05-04 introduction.tex

\section{Introduction and main results}

In this paper, we will consider clustering in a model of random graphs that involves points taken randomly in the hyperbolic plane. This model was introduced by Krioukov, Papadopoulos, Kitsak, Vahdat and Bogu\~{n}\'a~\cite{krioukov2010hyperbolic} in 
2010 - we abbreviate it as \emph{the KPKVB model}. We should however note that the model also goes by several other names in the literature, including {\em hyperbolic random geometric graphs} and {\em random hyperbolic graphs}. Krioukov et al.~suggested this model as a suitable model for complex networks. It exhibits the three main characteristics usually associated with complex networks: a power-law degree distribution, a non-vanishing clustering coefficient and short graph distances.

\subsection{KPKVB model}
We start with the definition of the model. As mentioned, its nodes are situated in the hyperbolic plane $\Haa$, which is a surface with constant Gaussian curvature $-1$. This surface has several convenient representations (i.e.~coordinate maps), such as the Poincar\'e half-plane model, the Poincar\'e disk model and the Klein disk model. A gentle introduction to Gaussian curvature, hyperbolic geometry and these representations of the hyperbolic plane can be found in~\cite{stillwell2012geometry}. Throughout this paper we will be working with a representation of the hyperbolic plane using {\em hyperbolic polar coordinates}, sometimes called the {\em native representation}. That is, a point $u \in \Haa$ is represented as $(r,\theta)$, where $r$ is the hyperbolic distance between $u$ and the origin $O$ and $\theta$ as the angle between the line segment $Ou$ and the positive $x$-axis. 
Here, when mentioning ``the origin'' and the angle between the line segment and the positive $x$-axis, we think of $\Haa$ embedded as the Poincar\'e disk in the ordinary euclidean plane.

The KPKVB model has three parameters: the number of vertices $n$, which we think of as going to infinity, and $\alpha > \frac{1}{2}$, $\nu > 0$ which we think of as fixed. Given $n, \alpha, \nu$ we define $R = 2\log(n/\nu)$. Then the hyperbolic random graph $G(n;\alpha, \nu)$ is defined as follows:
\begin{itemize}
\item The vertex set is given by $n$ i.i.d.~points $u_1, \dots, u_n$ denoted in polar coordinates $u_i = (r_i, \theta_i)$, where the angular coordinate $\theta$ is chosen uniformly from $(-\pi,\pi]$ while the radial coordinate $r$ is sampled independently according to the cumulative distribution function
\begin{equation}\label{eq:def_hyperbolic_point_distribution}
	F_{\alpha,R}(r) = \begin{cases}
		0 &\mbox{if } r < 0\\
		\frac{\cosh(\alpha r)-1}{\cosh(\alpha R) - 1} &\mbox{if } 0 \le r \le R\\
		1&\mbox{if } r > R
	\end{cases}
\end{equation}
\item Any two vertices $u_i=(r_i,\theta_i)$ and $u_j=(r_j,\theta_j)$ are adjacent if and only if $d_\H(u_i,u_j) \le R$, where $d_\H$ denotes the distance in the hyperbolic plane. We will frequently be using that, by the hyperbolic law of cosines, $d_\H(u_i,u_j) \le R$ is equivalent to
\[
	\cosh(r_i) \cosh(r_j) - \sinh(r_i) \sinh( r_j) \cos(|\theta_i-\theta_j|_{2\pi}) \le \cosh(R),
\]
where $|a|_{b} = \min( |a|, b - |a|)$ for $-b\leq a\leq b$.
\end{itemize}

\begin{figure}[!t]
\centering
\includegraphics[scale=0.3]{figures/KPKVB.png}
\caption{Simulation $G(n;\alpha, \nu)$ with $\alpha = 0.9$, $\nu = 0.2$ and $n = 5000$.}
\label{fig:H_graph_example}
\end{figure}

\noindent
Figure~\ref{fig:H_graph_example} shows a computer simulation of $G(n;\alpha, \nu)$.

As observed by Krioukov et al.~\cite{krioukov2010hyperbolic}, and proved rigorously by Gugelmann et al.~\cite{gugelmann2012random}, the 
degree sequence of the KPKVB model follows a power-law with exponent $2\alpha+1$.
Gugelmann et al.~\cite{gugelmann2012random} also showed that the average degree converges in probability to the 
constant $8\nu\alpha^2/ \pi (2\alpha-1)^2$, and they showed that the (local) clustering coefficient is non-vanishing 
in the sense that it is bounded below by a positive constant a.a.s. 
Here, and in the rest of the paper, for a sequence $(E_n)_n$ of events, $E_n$ {\em asymptotically almost surely} (a.a.s.) means 
that $\Prob{E_n} \to 1$ as $n \to \infty$.

Apart from the degree sequence and clustering, the third main characteristic associated with complex networks, ``short distances'', has 
also been established in the literature. In~\cite{abdullah2017typical} it is shown that for $\alpha < 1$ the largest component 
is what is called an \emph{ultra-small world}: if we randomly sample two vertices of the graph then, a.a.s., conditional on them 
being in the same component, their graph distance is of order $\log\log n$. 
In~\cite{kiwi2015bound} and~\cite{friedrich2018diameter} a.a.s.~polylogarithmic upper and lower bounds on the graph diameter of 
the largest component are shown, and in~\cite{muller2017diameter}, these were sharpened to show that $\log n$ is the correct order of 
the diameter.

Earlier work of the first and third authors with Bode~\cite{bode2015largest} and of the first and third authors~\cite{fountoulakis2018law} has established the ``threshold for a giant component'': if $\alpha < 1$ then there is a unique component of size linear in $n$ no matter how small $\nu$ (i.e. the average degree); if $\alpha > 1$ all components are sublinear no matter the value of $\nu$; and if $\alpha=1$ then there is a critical value $\nu_{\text{c}}$ such that for $\nu < \nu_{\text{c}}$ all components are sublinear and for $\nu > \nu_{\text{c}}$ there is a unique linearly sized component (all of these statements holding a.a.s.). Whether or not there is a giant component if $\alpha=1$ and $\nu=\nu_{\text{c}}$ remains an open problem. In~\cite{kiwi2015bound} and~\cite{kiwi2017second}, Kiwi and Mitsche considered the size of the second largest component and showed that for $\alpha \in (1/2, 1)$, a.a.s., the second largest component has polylogarithmic order with exponent $1/(\alpha -1/2)$.

In another paper of the first and third authors with Bode~\cite{bode2016probability} it was shown that $\alpha=1/2$ is the threshold for connectivity: for $\alpha < 1/2$ the graph is a.a.s.~connected, for $\alpha>1/2$ the graph is a.a.s.~disconnected and when $\alpha=1/2$ the probability of being connected tends to a continuous, non-decreasing function of $\nu$ which is identically one for $\nu \geq \pi$ and strictly less than one for $\nu < \pi$. Friedrich and Krohmer~\cite{blasius2018cliques} studied the size of the largest clique as well as the number of cliques of a given size. Bogu\~{n}a et al.~\cite{boguna2010sustaining} and Bl\"asius et al.~\cite{blasius2018efficient} considered fitting the KPKVB model to data using maximum likelihood estimation. Kiwi and Mitsche~\cite{kiwi2018spectral} studied the spectral gap and related properties, and Bl\"asius et al.\cite{blasius2016hyperbolic} considered the tree-width and related parameters of the KPKVB model. Recently Owada and Yogeshwaran~\cite{owada2018sub} considered subgraph counts, and in particular established a central limit theorem for the number of copies of a fixed tree $T$ in $G(n;\alpha,\nu)$, subject to some restrictions on the parameter $\alpha$.

\subsubsection*{Clustering}

In this work we study the clustering coefficient in the KPKVB model. In the literature there are unfortunately two distinct, rival definitions of the {\em clustering coefficient}. One of those, sometimes called the {\em global} clustering coefficient, is defined as three times the ratio of the number of triangles to the number of paths of length two in the graph. Results for this version of the clustering coefficient in the KPKVB model were obtained by Candellero and the first author~\cite{candellero2016clustering} and for the evolution of graphs on more general spaces with negative curvature by the first author in~\cite{fountoulakis2012evolution}.

We will study the other notion of clustering, the one which is also considered by Krioukov et al.~\cite{krioukov2010hyperbolic} and Gugelmann et al.~\cite{gugelmann2012random}. It is sometimes called the {\em local} clustering coefficient, although we should point out that Gugelmann et al.~actually call it the global clustering coefficient in their paper. For a graph $G$ and a vertex $v\in V(G)$ we define the clustering coefficient {\em of $v$} as:
\[
	c(v) := \left\{\begin{array}{cl}
		\displaystyle \frac{1}{\binom{\text{deg}(v)}{2}} \sum_{u,w\sim v} 1_{\{uw \in E(G)\}}, 
			& \text{ if $\text{deg}(v) \geq 2$, }\\
		& \\
        0, & \text{ otherwise,}
        \end{array}\right.
\]
where $E(G)$ denotes the edge set of $G$ and $\text{deg}(v)$ is the degree of vertex $v$. That is, provided $v$ has degree at least two, $c(v)$ equals the number of edges that are actually present between the neighbours of $v$ divided by the number of edges that could possibly be present between the neighbours given the degree of $v$.
The clustering coefficient of $G$ is now defined as the average of $c(v)$ over all vertices $v$:
\[
	c(G) := \frac{1}{|V(G)|} \sum_{v\in V(G)} c(v).
\]

As mentioned above, Gugelmann et al.~\cite{gugelmann2012random}, have established that $c(G(n;\alpha,\nu))$ is non-vanishing a.a.s., but they left open the question of convergence. Theorem~\ref{thm:maincc} below establishes that the clustering coefficient indeed converges in probability to a constant $\gamma$ that we give explicitly as a closed form expression involving $\alpha,\nu$ and several classical special functions.

In addition to the clustering coefficient, we shall also be interested in the {\em clustering function}.
This assigns to each non-negative integer $k$ the value
\begin{equation}\label{eq:def_clustering_function}
	c(k; G) := \begin{cases}
		\displaystyle \frac{1}{N(k)} \sum_{\substack{v \in V(G), \\ \text{deg}(v)=k}}  c(v),  &\mbox{ if } N(k) \ge 1,\\
		0, &\mbox{else,}
	\end{cases}
\end{equation}
where $N(k)$ denotes the number of vertices of degree exactly $k$ in $G$. In other words, the clustering function assigns to the integer $k$ the average of the local clustering coefficient over all vertices of degree $k$. We remark that, while it might seem natural to consider $c(k)$ to be ``undefined'' when $N(k)=0$, we prefer to use the above definition for technical 
convenience.  This way $c(k; G(n;\alpha,\nu) )$ is a plain vanilla random variable and we can for instance compute its moments without any issues.

Krioukov et al.~state (\cite{krioukov2010hyperbolic}, last sentence on page 036106-10) that as $k$ tends to infinity, the clustering function decays as $k^{-1}$. This seems to be based on computations that were not included in the paper.
Despite the attention the KPKVB model has generated since then, the behaviour of the clustering function in KPKVB random graphs has not been rigorously determined yet. In particular it has not been established whether it converges as $n\to\infty$ to some suitable limit function.  
Theorems~\ref{thm:mainkfixed} and~\ref{thm:mainktoinfty} below settle this question. Theorem~\ref{thm:mainkfixed} shows that for each fixed $k$, the value $c(k;G(n;\alpha,\nu))$ converges in probability to a constant $\gamma(k)$ that we again give explicitly as a closed form expression involving $\alpha,\nu$ and several classical special functions. Theorem~\ref{thm:mainktoinfty} extends this result to growing sequences satisfying $k \ll n^{1/(2\alpha+1)}$. Proposition~\ref{prop:asymp} clarifies the asymptotic behavior of the limiting function $\gamma(k)$, as $k\to\infty$. This depends on the parameter $\alpha$, and $\gamma(k)$ only scales with $k^{-1}$ when $\alpha > 3/4$, which corresponds to the exponent of the degree distribution exceeding $5/2$. 
So in particular our findings contradict the above-mentioned claim of Krioukov et al.~\cite{krioukov2010hyperbolic}.

\subsubsection*{Notation}

In the statement of our main results, and throughout the rest of the paper, we will use the following notations. 
We set 
$$\xi := \frac{4\alpha\nu}{\pi(2\alpha-1)}. $$

We write $\Gamma(z) := \int_0^\infty t^{z-1} e^{-t}\text{d}t$ for the gamma function, 
$\Gamma^+(a,b) := \int_b^\infty t^{a-1} e^{-t}\text{d}t$ for the upper incomplete gamma function, 
 $B(a,b) := \int_0^1 u^{a-1}(1-u)^{b-1}\text{d}u = \Gamma(a)\Gamma(b) / \Gamma(a+b)$ for the beta function and 
 $B^-(x ; a,b) := \int_0^x u^{a-1}(1-u)^{b-1}\text{d}u$ for the lower incomplete beta function. 
We write $U(a,b,z)$ for the hypergeometric U-function (also called Tricomi's confluent hypergeometric function), which 
%for $a,b,z\in \mathbb{C}$, $b \not \in \mathbb{Z}_{\leq 0}$, $\mathrm{Re}(a), \mathrm{Re}(z) >0$ 
has the integral representation 
\[
	U(a,b,z) = \frac{1}{\Gamma(a)} \int_0^\infty e^{-zt} t^{a-1} (1+t)^{b-a-1} dt,
\] 
see~\cite[p.255 Equation (2)]{erdelyi1953higher}, and let $\MeijerGnew{m}{\ell}{p}{q}{{\bf a}}{{\bf b}}{z}$ denote 
Meijer's G-Function~\cite{meijer1946gfunction}, see Appendix~\ref{sec:Meijer_G_functions} for more details.

For a sequence $(X_n)_n$ of random variables, we write $X_n \xrightarrow[n\to\infty]{\Pee} X$ to denote that $X_n$ converges in 
probability to $X$. %, and we write $X_n \xrightarrow[n\to\infty]{L_1} X$ if $X_n$ converges to $X$ in the $L_1$ sense (i.e.~if $\Ee |X_n-X| \to 0$.)

%We further adopt standard notations on asymptotic behavior of functions and sequences. That is, for two functions $f$ and $g$, we write $f(n) = \smallO{g(n)}$ as $n \to \infty$ if $\limsup_{n \to \infty} f(n)/g(n) = 0$ and $f(n) = \bigO{g(n)}$ as $n \to \infty$ if $\limsup_{n \to \infty} |f(n)|/g(n) < \infty$. Furthermore, $f(n) = \Omega(g(n))$ as $n \to \infty$ if $\limsup_{n \to \infty} |f(n)/g(n)| > 0$ and $f(n) = \omega(g(n))$ as $n \to \infty$ if $\limsup_{n \to \infty} |f(n)/g(n)| = \infty$. Finally, $f(n) = \bigT{g(n)}$ as $n \to \infty$ if $f(n) = \bigO{g(n)}$ and $g(n) = \Omega(f(n))$.

 %In addition, $X_n \xrightarrow[n\to\infty]{L^1} X$ denotes convergence in expectation, i.e. $\Exp{|X_n - X|} \to 0$ as $n \to \infty$, which is a stronger notion than convergence in probability.

\subsection{Main results}\label{ssec:main_results}

\subsubsection{The clustering coefficient}

Our first main result shows the convergence of the local clustering coefficient in the KPKVB model and 
establishes the limit exactly.

\begin{theorem}\label{thm:clustering_coefficient_hyperbolic}\label{thm:maincc}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ be fixed. Writing $G_n := G(n;\alpha,\nu)$, we have
\[
	c( G_n ) \xrightarrow[n\to\infty]{\Pee} \gamma,
\]
where $\gamma$ is defined for $\alpha \ne 1$ as
\begin{align*}
	\gamma 
	&=\frac{2 + 4 \alpha + 13 \alpha^2 - 34 \alpha^3 - 12\alpha^4 + 24 \alpha^5}{16(\alpha-1)^2 \alpha (\alpha+1) (2\alpha+1)} 
		+  \frac{2^{-1 - 4 \alpha}}{(\alpha - 1)^2} \\
&\hspace{10pt}+ \frac{(\alpha - 1/2) (B(2 \alpha, 2 \alpha + 1) + B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha))}{2 (\alpha - 1) (3 \alpha - 1)} \\
%
&\hspace{10pt}+ \frac{\xi^{2\alpha} \left( \Gamma^+( 1 - 2 \alpha, \xi) + \Gamma^+( - 2 \alpha, \xi)\right) }{4(\alpha-1)} \\
%
&\hspace{10pt}+ \frac{\xi^{2\alpha + 2}\alpha (\alpha - 1/2)^2 \left( \Gamma^+(- 2 \alpha - 1, \xi) + \Gamma^+(- 2 \alpha - 2, \xi)\right)}%
{2(\alpha-1)^2} \\
%
&\hspace{10pt}- \frac{\xi^{2\alpha + 1}\alpha (2\alpha - 1) \left( \Gamma^+( - 2\alpha,\xi)+\Gamma^+( - 2 \alpha - 1,\xi) \right)}%
{(\alpha-1)} \\
%
&\hspace{10pt}- \frac{\xi^{6\alpha-2}2^{-4\alpha}(3\alpha - 1)
\left( \Gamma^+( - 6 \alpha + 3, \xi)+\Gamma^+( - 6 \alpha + 2, \xi) \right)}{(\alpha-1)^2}\\
%
&\hspace{10pt}- \frac{\xi^{6\alpha - 2}(\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha)%
\left(\Gamma^+( - 6 \alpha + 3, \xi)+\Gamma^+( - 6 \alpha + 2, \xi)\right)}{(\alpha-1)} \\
%
&\hspace{10pt}- \frac{e^{-\xi} \Gamma(2\alpha+1) \left(U(2\alpha+1,1-2\alpha,\xi) + U(2\alpha+1,2-2\alpha,\xi)\right)}{4(\alpha-1)} \\
%
&\hspace{10pt}+ \frac{\xi^{6\alpha - 2} \Gamma(2\alpha+1)\left( \MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+2,0}{\xi}
 		+ \MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+3,0}{\xi}\right)}{4(\alpha-1)},
\end{align*}
and for $\alpha = 1$ as the $\alpha\to 1$ limit of the above expression. 
% \begin{align*}
% 	c_\infty &= \frac{575 - 12 \pi^2}{576} + \frac{\eta^4(7 + \pi^2)\Gamma^\ast(-4, \eta)}{4}\\
% 	&\hspace{10pt}- \frac{1}{2} \int_0^1 (1 - 4z + 3z^3)\log(1-z)(z + \eta)e^{-\eta/z} \dd z\\
% 	&\hspace{10pt}- \int_0^1 \Li_2(z)(z^3 + \eta z^2) e^{-\eta/z} \dd z,		
% \end{align*}
% with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function\footnote{Note that the integrals in the expression for $c_\infty$ for $\alpha = 1$ exists: for the first one note that $1-4z+3z^2=(1-z)(1-3z)$, so the integrand can be bounded by $C(1-z)\log(1-z)$ on $[0,1)$ for some constant $C$, which can be continued continuously to the compact interval $[0,1]$ by noting that the limit for $z \rightarrow 1$ is zero, so the integrand is bounded on a bounded domain and hence, this integral is finite; for the second integral note that $\Li_2(z)$ is bounded by $\Li_2(1)$ on $[0,1]$, which is a series with well-known finite limit, so again the integrand is bounded on a bounded domain and hence the second integral is also finite.}.
\end{theorem}

\noindent
A plot of $\gamma$ can be found in Figure~\ref{fig:gamma}. The figure also shows the results of computer simulations that appear to be in agreement with our findings.

\begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.6]{figures/cn10000nu0512rep100a05to5Squares.pdf}
    \caption{Plot of $\gamma$ for $\alpha$ varying from $0.5$ to $5$ on the horizontal axis and 
    for $\nu=\frac{1}{2}$ (blue), $\nu=1$ (purple), $\nu=2$ (green). Simulations (squares in corresponding colour) with $n=10000$ and $100$ repetitions.\label{fig:gamma}}
\end{figure}

In the above expression for $\gamma$, a factor $\alpha-1$ occurs in the denominator of each term, but we will see that this
corresponds to a removable singularity. We have not been able to find a closed form expression in terms of standard functions in the case 
when $\alpha=1$, but in Section~\ref{ssec:alphais1} we do provide an explicit expression involving integrals.

\subsubsection{The clustering function}

Our second main result is on the clustering function for constant $k$.

\begin{theorem}\label{thm:local_clustering_hyperbolic}\label{thm:mainkfixed}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $k\geq2$ be fixed. 
Writing $G_n := G(n;\alpha,\nu)$, we have

\[
	c(k;G_n) \xrightarrow[n\to\infty]{\Pee} \gamma(k),
\]
where $\gamma(k)$ is defined for $\alpha \ne 1$ as 
\begin{align*}
\gamma(k)  =&\frac{1}{8\alpha (\alpha-1)\Gamma^+(k-2\alpha,\xi)} \left( -\Gamma^+(k - 2 \alpha, \xi) - 2\frac{\alpha (\alpha - 1/2)^2 \xi^{2} \Gamma^+(k - 2 \alpha - 2, \xi)}{(\alpha - 1)} \right. \\ 
&\left.+ 8 \alpha (\alpha - 1/2) \xi \Gamma^+(k - 2 \alpha - 1,\xi) \right.\\ 
&\left.+ 4\xi^{4\alpha - 2} \Gamma^+(k - 6 \alpha + 2, 
      \xi) \left( \frac{2^{ - 4\alpha}(3 \alpha - 1)}{(\alpha - 1)} + (\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha) \right)  \right.\\ 
&\left.+ \xi^{k-2\alpha} \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi) \right. \\ 
&\left.- \xi^{4\alpha-2} \Gamma(2\alpha+1)\MeijerGnew{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}  \right)
\end{align*}
and for $\alpha = 1$ as the $\alpha\to1$ limit of the above expression.
\end{theorem}

\noindent
A plot of $\gamma(k)$,  together with the results of computer experiments, can be found in Figure~\ref{fig:gammak}. %
%
%
\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.6]{figures/ckn10000a08nu0512rep100k2to25Squares.pdf}
    \caption{Plot $\gamma(k)$ for $k$ varying from 2 to 25 on the horizontal axis, for $\alpha=0.8$ and $\nu=\frac{1}{2}$ 
    (blue), $\nu=1$ (purple), $\nu=2$ (green). Simulations (squares in corresponding colour) with $n=10000$ and $100$ repetitions.\label{fig:gammak}}
\end{figure}%
%
Again, we remark that the above expression for $\gamma(k)$ appears to have a singularity at $\alpha=1$, but this will turn out to be a removable singularity. 
Again, we have not been able to find a closed form expression in terms of standard functions in the case when $\alpha=1$, but in 
Section~\ref{ssec:alphais1} we do provide an explicit expression involving integrals.

Theorem~\ref{thm:mainkfixed} in fact generalises to increasing sequences $(k_n)_{n \ge 1}$.

\begin{theorem}\label{thm:mainktoinfty}
Let $\alpha>\frac12, \nu>0$ be fixed and let $k_n$ be a sequence of non-negative integers
satisfying $1 \ll k_n \ll n^{1/(2\alpha+1)}$. Then, writing $G_n := G(n;\alpha,\nu)$, we have

$$ \frac{c(k_n;G_n)}{\gamma(k_n)} \xrightarrow[n\to\infty]{\Pee} 1. $$

\end{theorem}

\noindent
The statement of Theorem~\ref{thm:mainktoinfty} is equivalent to $c(k_n;G_n) = (1+o(1)) \gamma(k_n)$ a.a.s., using notation that is common in the random graphs community. %$c(k_n; G_n)/\gamma(k_n)\xrightarrow[n\to\infty]{\Pee} 1$, which might alternatively be written as

\subsubsection{Scaling of \texorpdfstring{$\gamma(k)$}{gamma(k)}}

To clarify the scaling behaviour of $\gamma(k)$ we offer the following result.

\begin{proposition}\label{prop:asymp}
As $k\to\infty$, we have

$$ \gamma(k) = 
\left\{ \begin{array}{ll}
(c_{\alpha,\nu}+o(1)) \cdot k^{-1} &\text{ if } \alpha > \frac{3}{4}, \\
(c_{\alpha,\nu}+o(1)) \cdot \frac{\log(k)}{k}& \text{ if } \alpha = \frac{3}{4},\\
(c_{\alpha,\nu}+o(1)) \cdot k^{2-4\alpha} & \text{ if } \frac12 < \alpha < \frac34, 
\end{array} \right.,
$$
where

$$ c_{\alpha,\nu} := 
\left\{ \begin{array}{cl}
8\alpha \nu / (\pi\left(4\alpha - 3\right)) & \text{ if } \alpha > \frac{3}{4}, \\
6 \nu / \pi & \text{ if } \alpha = \frac{3}{4},\\
 \left( \frac{3\alpha - 1}{2^{4\alpha+1}\alpha(\alpha-1)^2} 
	+ \frac{(\alpha - \frac{1}{2})B^-(\frac{1}{2},2\alpha + 1, 2\alpha - 2)}{2(\alpha - 1)\alpha} 
	- \frac{B(2\alpha, 3\alpha - 4)}{4(\alpha - 1)} \right)  \cdot \xi^{4\alpha-2} 
	& \text{ if } \frac12 < \alpha < \frac34.
\end{array} \right. $$
\end{proposition}

Theorem~\ref{thm:mainktoinfty} states that the clustering function of the KPKVB model scales as $\gamma(k)$ as the number of vertices $n\to\infty$, and
Proposition~\ref{prop:asymp} makes clear how $\gamma(k)$ behaves as $k$ grows.
In particular, these results contradict the scaling claimed in~\cite{krioukov2010hyperbolic} for 
$\alpha \leq \frac{3}{4}$, and confirms it only for $\alpha > \frac{3}{4}$.

We remark that simultaneously and independently Stegehuis, van der Hofstad and van Leeuwaarden~\cite{stegehuis2019scale} used a completely different technique to obtain a similar, though less detailed, result on the $k\to\infty$ scaling of the clustering function in the KPKVB model.

\subsection{Additional observations and results}

There are a few additional remarks we would like to make regarding our results.

\subsubsection{The degree distribution and the range of \texorpdfstring{$k_n$}{kn} in Theorem~\ref{thm:mainktoinfty}}

The reader may already have observed that, with a power law exponent of $2\alpha+1$ for the probability mass function of 
the degree sequence, we would expect $\Theta( n \cdot k^{-(1+2\alpha)} ) = o(1)$ vertices of degree {\em exactly} $k$ whenever
$k \gg n^{1/(1+2\alpha)}$.

This is the reason why in Theorem~\ref{thm:mainktoinfty} we restrict ourselves to 
sequences $k_n$ with $k_n \ll n^{1/(1+2\alpha)}$. 
When  $k_n \gg n^{1/(1+2\alpha)}$ there are no vertices of degree exactly $k_n$ a.a.s., which in particular
implies that the clustering function equals zero a.a.s.~for any such sequence $k_n$.
%When $k_n$ is of the same order as $n^{1/(1+2\alpha)}$ then the probability that there are no vertices of degree exactly $k_n$
%stays bounded away from zero.

As mentioned previously, Gugelmann et al.~\cite{gugelmann2012random} gave a mathematically rigorous result on the degree sequence, which can of course be rephrased as a result on the number of nodes with degree exactly $k$. 
Their results allow $k = k_n$ to grow with $n$, but unfortunately require that $k_n \leq n^\delta$ with 
$\delta < \min\left\{\frac{2\alpha-1}{4\alpha(2\alpha +1)}, \frac{2(2\alpha -1)}{5(2\alpha + 1)}\right\} < \frac{1}{2\alpha + 1}$.
For completeness we offer the following result, which extends that of Gugelmann et al.~to the full range $1 \le k_n \le n-1$.

\begin{theorem}\label{thm:degrees_hyperbolic}
Let $\alpha >\frac{1}{2}, \nu > 0$, denote by $N_n(k)$ the number of vertices with degree $k$ in the KPKVB model $G(n;\alpha,\nu)$ and 
consider a sequence of integers $(k_n)_n$ with $0 \leq k_n \leq n-1$.
\begin{enumerate}
\item If $k_n \ll n^{\frac{1}{2\alpha+1}} $ as $n\rightarrow \infty$, then a.a.s.

$$N_n(k_n) = (1+o(1)) \cdot n \cdot \pmf(k_n),$$

where $\pmf(k_n) = 2\alpha \xi^{2\alpha}\Gamma^+(k_n-2\alpha,\xi)/k_n!$.

\item If $k_n = (1+o(1))c n^{\frac{1}{2\alpha+1}}$ for some fixed $c >0$, then

$$ N_n(k_n) \xrightarrow[n\rightarrow\infty]{d} \operatorname{Po}(2\alpha \xi^{2\alpha} c^{-(2\alpha+1)}). $$

\item If $k_n \gg n^{\frac{1}{2\alpha+1}}$, then a.a.s. $N_n(k_n) = 0$.
\end{enumerate}
\end{theorem}

\subsubsection{Transition in scaling at \texorpdfstring{$\alpha = 3/4$}{alpha equals 3/4}.}

Proposition~\ref{prop:asymp} demonstrates that there is a transition in the scaling of the local clustering 
function at $\alpha = 3/4$. 
This corresponds to an exponent $5/2$ for the probability mass function of the degree distribution. 
This transition is different from those often observed for networks with scale-free degree distributions, where transitions occur at
integer values of the exponent. 
At this point, it is unclear what the underlying reason is for the appearance of the transition at this particular half integer 
exponent. 
Interestingly, a similar transition point has also been observed for both majority vote models~\cite{chen2015critical} and 
flocking dynamics~\cite{miguel2018effects} on networks with scale-free degree degree distributions.

%\subsubsection{Uniform convergence.}

%Our results for the local clustering function in fact imply uniform convergence of $c(k; G_n)$ for all $2 \le k \le a_n$ 
%where $a_n \ll n^{\frac{1}{2\alpha + 1}}$. 
%To see this we first observe that for {\em fixed} $k$, the statement $c(k;G_n) \xrightarrow[n\to\infty]{\Pee} \gamma(k)$
%is equivalent to $c(k;G_n)/\gamma(k) \xrightarrow[n\to\infty]{L_1} 1$.
%Now let
%\[
%	b_n = \arg \max_{2 \le k \le a_n} \Exp{\left|\frac{c(k; G_n)}{\gamma(k)} - 1\right|}.
%\]
%Then $b_n \le a_n \ll n^{\frac{1}{2\alpha + 1}}$ and therefore by Theorem~\ref{thm:mainkfixed} and Theorem~\ref{thm:mainktoinfty}
%\[
%	\lim_{n \to \infty} \max_{2 \le k \le a_n} \Exp{\left|\frac{c(k; G_n)}{\gamma(k)} - 1\right|} 
%	= \lim_{n \to \infty}\Exp{\left|\frac{c(b_n; G_n)}{\gamma(b_n)} - 1\right|} = 0.
%\]

\subsection{Outline of the paper}

In the next section we will recall some useful tools from the literature and define a series of auxiliary random graph models 
that will be used in the proofs. In particular, we relate in a series of steps the KPKVB model to an infinite percolation model 
$\Ginf$ that was used in previous work of the first and third authors~\cite{fountoulakis2018law} on the largest component 
of the KPKVB model. 
The value of the limiting constant $\gamma$, respectively limiting clustering function $\gamma(k)$, correspond to the probability 
that two randomly chosen neighbours of a ``typical point'' in this infinite model are themselves neighbours, respectively the 
probability of this event conditional on the typical point having exactly $k$ neighbours. These probabilities can be expressed as 
certain integrals, which we solve explicitly in Section~\ref{sec:Ginf}. 
In the same section we also prove Proposition~\ref{prop:asymp}, on the asymptotics of $\gamma(k)$. 
We then proceed to prove Theorems~\ref{thm:maincc} and~\ref{thm:mainkfixed} by relating said probabilities for the typical point 
of the infinite model to the corresponding clustering coefficient/function in the original KPKVB random graph, using the Campbell-Mecke 
formula and some other, relatively straightforward considerations.

In Section~\ref{sec:degrees} we prove Theorem~\ref{thm:degrees_hyperbolic}, which also doubles as a warm-up for the proof 
of Theorem~\ref{thm:mainktoinfty} of the clustering function for growing $k$.
The remaining sections are devoted to the proof of Theorem~\ref{thm:mainktoinfty}, which turns out to be technically involved.
The main reason for this is that when we push $k_n$ close to the maximum possible value, a great deal of work is needed 
to properly control the arising error terms.

The Appendix includes some auxiliary results on Meijer's G-function, Chernoff bounds for Poisson and Binomial 
random variables and the code used for simulations.